---
title: "Bias in EPSRC Peer Review - Research Question 5"
author: 'Contributors: Ged Hiscoke & Dakota Langhals, Royal Statistical Society'
date: "Version: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    code_folding: hide
    toc_float: yes
  pdf_document: default
  word_document:
    reference_docx: "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/r_style_template.docx"
    highlight: null
subtitle: 'Status: Final'
---

```{r setup, include=FALSE}

##Universal Document settings
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12, warning = FALSE, message = FALSE)


## Libraries
library(kableExtra) # For tables
library(ggplot2)    # For plotting
library(patchwork)  # For plot displays
library(tidyr)      # For tidy coding
library(ggforce)    # For tidy labelling on plots 
library(gridExtra)  # For display of joint plots
library(knitr)
library(openxlsx)   # for saving tables as excel files
library(scales)     # For commas in plot labels
library(broom)
library(lme4)
library(broom)
library(pscl)  # For pR2 function
library(pROC)  # For ROC AUC
library(ModelMetrics)  # For Brier Score
library(DescTools) 
library(tidyverse)
library(boot)
library(ggrepel)
library(dplyr)
library(performance)
library(margins)





## Set key file locations
## Directory for saving local high resolution copies of plots etc
data_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Datasets/"
note_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Notes/"
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05"
table_folder_path   <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Tables/Note05"



## Set key file names
raw_data_file       <- "PROJ2023.002-DATA001.xlsx"
BaselineEnvironment <- "PROJ2023.002 Environment - BASELINE.RData"
CleanedEnvironment  <- "PROJ2023.002 Environment - CLEANED.RData"
Sample20Environment <- "PROJ2023.002 Environment - SAMPLE20.RData"
Sample80Environment <- "PROJ2023.002 Environment - SAMPLE80.RData"
cleanedExcelFile    <- "PROJ2023.002-DATA002.xlsx"
WorkingEnvironment  <- "PROJ2023.002 Environment - WORKING.RData"
UKLFdata            <- "PROJ2023.002-DATA004.xlsx"


## Load needed data
load(file = file.path(data_folder_path, WorkingEnvironment))



## Define the RSS colour palette

RRSSgold      <- "#D3A41B" 
RSSblue       <- "#004B6D" 
RSSred        <- "#cc0000" 
RSStaupe      <- "#a6a086" 
RSSgreen      <- "#009933" 
RSSpurple     <- "#7879b9" 
RSSdark       <- "#393f4d" 
RSSlightblue    <- "#6db4c0" 
RSSseablue       <- "#287fc3" 
RSSorange     <- "#d86027"
RSSlightgrey  <- "#bdbdbd"
RSSmidgrey    <- "#939393"
RSSdarkkgrey  <- "#7e7e7e"
RSSjungle     <- "#1f3d0c"
RSSmint       <- "#5C9090" 
RSSherb       <- "#689E88"
RSSbright     <- "#8D31A4"
RSSlime       <- "#7DCF7A"
RSSviolet     <- "#614182"


##Variable specific colours
col_male      <- "#004b6d"
col_female    <- "#a6700a"
col_uk        <- "#393f4d"
col_nonuk     <- "#614182"
col_dis       <- "#005b49"
col_nodis     <- "#8e8e00"
col_white     <- "#7f3c21"
col_ethmin    <- "#3a3877"
col_unknown   <- "#bdbdbd"
col_notdisclosed  <- "#640106"
col_age1 <- "#b6d6b3"
col_age2 <- "#87bd00"
col_age3 <- "#299907"
col_age4 <- "#146f03"
col_age5 <- "#054907"
col_age6 <- "#0f2808"
col_eth1 <- "#a7a7c1"
col_eth2 <- "#74749d"
col_eth3 <- "#1e3399"
col_eth4 <- "#090771"
col_eth5 <- "#37163e"
col_eth6 <- "#59095f"
col_eth7 <- "#34086e"
```


# {.tabset}

## Data Preparation

```{r}
## This dataframe will include Grant (rather than unique researcher)

## 1.1 Select columns from Grants.df
Grants_selected.df <- EPSRC$Grants.df %>%
  select(AnonPROJ,
         GrantCategory,
         GrantOutcome,
         AdjustedApplicationValue,
         Driver,
         AnonREG
  )


## 1.2 Now select additional columns from PrincipalInvestigators.df
PI_details.df <- EPSRC$PI.df %>%
  select(
    AnonPROJ,
    AnonINV,
    Sex,
    AgeRange,
    Disability,
    EthnicGroup,
    EthnicityBinary,
    NationalityBinary,
    AnonTHM,
    Role
  )


## 1.3 Merge with PrincipalInvestigators.df
Q5Grants.df <- left_join(PI_details.df,
                         Grants_selected.df ,
                         by = "AnonPROJ"
)


## 1.4 Cleanup
rm(PI_details.df, 
   Grants_selected.df
)

## Add a binary Outcome variable for Funded/Unfunded
Q5Grants.df$BinaryOutcome <- ifelse(Q5Grants.df$GrantOutcome == "Funded", "Funded", "Unfunded")


#Fundable by score
# Group by AnonPROJ and count the number of TRUE and FALSE FundableByScore values
result <- EPSRC$GrantMeeting.df %>%
  mutate(Fundablebyscore = FundableByScore)
  
result <- result %>% 
  group_by(AnonPROJ) %>%
  dplyr::summarise(
    TRUE_count = sum(Fundablebyscore == TRUE, na.rm = TRUE),
    FALSE_count = sum(Fundablebyscore == FALSE, na.rm = TRUE)
  ) %>%
  filter(FALSE_count == 0 & !is.na(FALSE_count))  # Keep rows where FALSE_count is 0 and not NA


## Add to Q5Grants.df
Q5Grants.df <- Q5Grants.df %>%
  left_join(result, by = "AnonPROJ") %>%
  mutate(Fundability = case_when(
    GrantOutcome == "Funded" ~ "Funded",
    !is.na(TRUE_count) & TRUE_count != 0 ~ "Fundable",
    TRUE ~ "Not Fundable"
  )) %>%
  select(-TRUE_count, -FALSE_count) 


# Collapsing age to align with ATI and make interactions more easily interpretable
Q5Grants.df <- Q5Grants.df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35 and Under", 
                                       AgeRange == "26-35" ~ "35 and Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56 and Over",
                                       AgeRange == "66+" ~  "56 and Over",
                                       AgeRange == "Unknown" ~ "Unknown"))

```


## Proportions of Demographic Groups That Are Funded {.tabset}

### Funded and Fundability Rates within Demographics, Separating Research Grants from Fellowships

```{r}
# Compare Pops ------------------------------------------------------------
# Function to compute confidence intervals for proportions for each level of Fundability
compute_fundability_ci <- function(df, num_bootstrap_replicates = 1000) {
  # Total count of rows in the dataframe
  total_count <- nrow(df)
  
  if (total_count == 0) {
  return(data.frame(Fundability = character(0), Count = numeric(0), Proportion = numeric(0), CI_L = numeric(0), CI_U = numeric(0)))
  }
  
  # Define bootstrap function to calculate the proportion of a specific Fundability level
  bootstrap_proportion <- function(data, indices, fundability_level) {
    d <- data[indices, ]
    return(sum(d$Fundability == fundability_level) / total_count)
  }
  
  # Initialize a list to store the results
  results <- list()
  
  # Levels of Fundability
  fundability_levels <- c("Funded", "Fundable", "Not Fundable")
  
  # Compute CIs for each Fundability level
  for (fundability in fundability_levels) {
    # Subset the dataframe for the current level of Fundability
    subset_df <- df %>% filter(Fundability == fundability)
    
    # Calculate the count and proportion
    count <- nrow(subset_df)
    proportion <- count / total_count
    
    # Compute bootstrap confidence intervals
    boot_obj <- boot(data = df, statistic = function(data, indices) {
      bootstrap_proportion(data, indices, fundability)
    }, R = num_bootstrap_replicates)
    
    ci <- boot.ci(boot_obj, type = "perc")
    results[[fundability]] <- data.frame(
      Fundability = fundability,
      Count = count,
      Proportion = proportion,
      CI_L = ci$percent[4],
      CI_U = ci$percent[5]
    )
  }
  
  # Combine the results into a single dataframe
  final_results <- bind_rows(results)
  
  return(final_results)
}

# Function to iterate compute_fundability_ci over each level of demo_column
compute_fundability_ci_by_demo <- function(df, demo_column, num_bootstrap_replicates = 1000) {
  # Get unique levels of the demographic column
  demo_levels <- unique(df[[demo_column]])
  
  # Initialize a list to store the results for each demo level
  all_results <- list()
  
  # Iterate over each level of the demographic column
  for (demo_level in demo_levels) {
    # Subset the dataframe for the current level of the demographic column
    subset_df <- df %>% filter(df[[demo_column]] == demo_level)
    
    # Compute the CIs for the current subset
    ci_results <- compute_fundability_ci(subset_df, num_bootstrap_replicates)
    
    # Add the demographic level to the results
    ci_results <- ci_results %>% mutate(Demographic = demo_level)
    
    # Store the results in the list
    all_results[[demo_level]] <- ci_results
  }
  
  # Combine the results into a single dataframe
  final_results <- bind_rows(all_results) %>%
    mutate(Group = demo_column)  # Add the Group column
  
  return(final_results)
}

df <- Q5Grants.df %>% filter(Role == "Principal Investigator")

df_sex_known <- df %>% filter(Sex != "Unknown", Sex != "Not Disclosed")
df_ethn_known <- df %>% filter(EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed")
df_nationality_known <- df %>% filter(NationalityBinary != "Unknown", NationalityBinary != "Not Disclosed")
df_disability_known <- df %>% filter(Disability != "Unknown", Disability != "Not Disclosed")
df_age_known <- df %>% filter(AgeRange != "Unknown", AgeRange != "Not Disclosed", !is.na(AgeRange))

t1 <- compute_fundability_ci_by_demo(df_sex_known, "Sex")
t1 <- rbind(t1, compute_fundability_ci_by_demo(df_ethn_known, "EthnicityBinary"))
t1 <- rbind(t1, compute_fundability_ci_by_demo(df_nationality_known, "NationalityBinary"))
t1 <- rbind(t1, compute_fundability_ci_by_demo(df_disability_known, "Disability"))
t1 <- rbind(t1, compute_fundability_ci_by_demo(df_age_known, "AgeRangeCollapsed"))

t1$Application_Type <- "Research Grants"

df <- Q5Grants.df %>% filter(Role == "Fellow")

df_sex_known <- df %>% filter(Sex != "Unknown", Sex != "Not Disclosed")
df_ethn_known <- df %>% filter(EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed")
df_nationality_known <- df %>% filter(NationalityBinary != "Unknown", NationalityBinary != "Not Disclosed")
df_disability_known <- df %>% filter(Disability != "Unknown", Disability != "Not Disclosed")
df_age_known <- df %>% filter(AgeRange != "Unknown", AgeRange != "Not Disclosed", !is.na(AgeRange))

t2 <- compute_fundability_ci_by_demo(df_sex_known, "Sex")
t2 <- rbind(t2, compute_fundability_ci_by_demo(df_ethn_known, "EthnicityBinary"))
t2 <- rbind(t2, compute_fundability_ci_by_demo(df_nationality_known, "NationalityBinary"))
t2 <- rbind(t2, compute_fundability_ci_by_demo(df_disability_known, "Disability"))
t2 <- rbind(t2, compute_fundability_ci_by_demo(df_age_known, "AgeRangeCollapsed"))

t2$Application_Type <- "Fellowships"

combined_table <- rbind(t1, t2)

combined_table <- combined_table %>% 
  mutate(Fundability = factor(Fundability, levels = c("Funded", "Fundable", "Not Fundable")))


```



```{r}
custom_colors <-
  c(
    "Male"= RSSblue,      
    "Female"= RSSred,    
    "UK"=RSSblue,        
    "Non UK"=RSSred,   
    "Undisclosed Nationality" = RSSmidgrey,
    "Known Disability"=RSSred   ,    
    "No Known Disability"=RSSblue  ,   
    "White"=RSSblue  ,   
    "Ethnic Minority"=RSSred  ,  
    "56 and Over" = col_5665,
    "36-55" = col_3645,
    "35 and Under" = col_2635
  )
```

#### Fundability by Sex with Confidence Intervals

```{r}
# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "Sex")

sex_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications, by PI Sex",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


sex_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/sex_fundability.png"), 
       plot = sex_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```


#### Fundability by Ethnicity with Confidence Intervals

```{r}
# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "EthnicityBinary")

ethnicity_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications, by PI Ethnicity",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


ethnicity_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/ethnicity_fundability.png"), 
       plot = ethnicity_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```



#### Fundability by Age with Confidence Intervals

```{r}
# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "AgeRangeCollapsed")

age_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications,\nby PI Age Group",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


age_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/age_fundability.png"), 
       plot = age_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```



#### Fundability by Nationality with Confidence Intervals

```{r}
# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "NationalityBinary")

nationality_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications,\nby PI Nationality",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


nationality_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/nationality_fundability.png"), 
       plot = nationality_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```




#### Fundability by Disability with Confidence Intervals

```{r}
# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "Disability")

disability_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications,\nby PI Disability",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


disability_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/disability_fundability.png"), 
       plot = disability_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```



##Intersections

```{r Intersectional Component}

## Research Grants
df <- Q5Grants.df %>% filter(Role == "Principal Investigator")

# Collapsing age 
df <- df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35 and Under", 
                                       AgeRange == "26-35" ~ "35 and Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56 and Over",
                                       AgeRange == "66+" ~  "56 and Over"))

# Creating dataframe for intersectional analysis
df_sex_ethnicity <- df %>% 
  filter(Sex != "Unknown", Sex != "Not Disclosed", EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed") %>%
  mutate(Sex_Ethnicity = interaction(Sex, EthnicityBinary)) %>% 
  select(AnonPROJ, Sex_Ethnicity, Fundability)

df_sex_age <- df %>% 
  filter(Sex != "Unknown", Sex != "Not Disclosed", AgeRangeCollapsed != "Unknown", AgeRangeCollapsed != "Not Disclosed", ) %>%
  mutate(Sex_Age = interaction(Sex, AgeRangeCollapsed)) %>% 
  select(AnonPROJ, Sex_Age, Fundability)

df_ethnicity_age <- df %>% 
  filter(EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed", AgeRangeCollapsed != "Unknown", AgeRangeCollapsed != "Not Disclosed") %>%
  mutate(Ethnicity_Age = interaction(EthnicityBinary, AgeRangeCollapsed)) %>% 
  select(AnonPROJ, Ethnicity_Age, Fundability)


t1 <- compute_fundability_ci_by_demo(df_sex_ethnicity, "Sex_Ethnicity")
t1 <- rbind(t1, compute_fundability_ci_by_demo(df_sex_age, "Sex_Age"))
t1 <- rbind(t1, compute_fundability_ci_by_demo(df_ethnicity_age, "Ethnicity_Age"))


t1$Application_Type <- "Research Grants"




## Fellowships
df <- Q5Grants.df %>% filter(Role == "Fellow")

# Collapsing age 
df <- df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35 and Under", 
                                       AgeRange == "26-35" ~ "35 and Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56 and Over",
                                       AgeRange == "66+" ~  "56 and Over"))

# Creating dataframe for intersectional analysis
df_sex_ethnicity <- df %>% 
  filter(Sex != "Unknown", Sex != "Not Disclosed", EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed") %>%
  mutate(Sex_Ethnicity = interaction(Sex, EthnicityBinary)) %>% 
  select(AnonPROJ, Sex_Ethnicity, Fundability)

df_sex_age <- df %>% 
  filter(Sex != "Unknown", Sex != "Not Disclosed", AgeRangeCollapsed != "Unknown", AgeRangeCollapsed != "Not Disclosed", ) %>%
  mutate(Sex_Age = interaction(Sex, AgeRangeCollapsed)) %>% 
  select(AnonPROJ, Sex_Age, Fundability)

df_ethnicity_age <- df %>% 
  filter(EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed", AgeRangeCollapsed != "Unknown", AgeRangeCollapsed != "Not Disclosed") %>%
  mutate(Ethnicity_Age = interaction(EthnicityBinary, AgeRangeCollapsed)) %>% 
  select(AnonPROJ, Ethnicity_Age, Fundability)


t2 <- compute_fundability_ci_by_demo(df_sex_ethnicity, "Sex_Ethnicity")
t2 <- rbind(t2, compute_fundability_ci_by_demo(df_sex_age, "Sex_Age"))
t2 <- rbind(t2, compute_fundability_ci_by_demo(df_ethnicity_age, "Ethnicity_Age"))


t2$Application_Type <- "Fellowships"


# # Filter for each interaction variable to ensure more than one row per level of Fundability (necessary for the function to work)
# df_filtered_sex_ethnicity <- df_intersections %>%
#   group_by(Sex_Ethnicity) %>%
#   filter(n_distinct(Fundability) > 1) %>%
#   ungroup()
# 
# df_filtered_sex_age <- df_intersections %>%
#   group_by(Sex_Age) %>%
#   filter(n_distinct(Fundability) > 1) %>%
#   ungroup()
# 
# df_filtered_ethnicity_age <- df_intersections %>%
#   group_by(Ethnicity_Age) %>%
#   filter(n_distinct(Fundability) > 1) %>%
#   ungroup()

```

```{r}
combined_table <- rbind(t1, t2)

combined_table <- combined_table %>% 
  mutate(Fundability = factor(Fundability, levels = c("Funded", "Fundable", "Not Fundable")))


```

### Sex + Ethnicity

```{r}
# Define custom colors
custom_colors <- c(
    "Male.White"= RSSblue,      
    "Male.Ethnic Minority"= RSSlightblue,    
    "Female.White"= RSSred   ,    
    "Female.Ethnic Minority"= RSSgold
  )

# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "Sex_Ethnicity")

sex_ethnicity_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications, by PI Sex*Ethnicity",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


sex_ethnicity_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/sex_ethnicity_fundability.png"), 
       plot = sex_ethnicity_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```



### Sex + Age

```{r}
# Define custom colors
custom_colors <- c(
    "Male.56 and Over"= RSSblue,      
    "Male.36-55"= RSSlightblue, 
    "Male.35 and Under"= RSSmint,  
    "Female.56 and Over"= RSSred   ,    
    "Female.36-55"= RSSorange,
    "Female.35 and Under"= RSSgold
  )

# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "Sex_Age")

sex_age_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications, by PI Sex*Age",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


sex_age_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/sex_age_fundability.png"), 
       plot = sex_age_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```



### Ethnicity + Age

```{r}
# Define custom colors
custom_colors <- c(
    "White.56 and Over"= RSSblue,      
    "White.36-55"= RSSlightblue, 
    "White.35 and Under"= RSSmint,  
    "Ethnic Minority.56 and Over"= RSSred   ,    
    "Ethnic Minority.36-55"= RSSorange,
    "Ethnic Minority.35 and Under"= RSSgold
  )

# Plot the proportions and confidence intervals

df <- combined_table %>% filter(Group == "Ethnicity_Age")

sex_age_fundability <- ggplot(df, aes(x = Fundability, y = Proportion, fill = Demographic)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), width = 0.2, position = position_dodge(0.9)) +
  scale_fill_manual(values = custom_colors) +
  theme_minimal() +
  labs(title = "Rates of Funded, Fundable, and Not Fundable Applications, by PI Ethnicity*Age",
       x = element_blank(),
       y = "Percentage") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    #axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
    ) +
  geom_text(aes(y = 0.03, label = scales::percent(Proportion, accuracy = 0.1)),
            size = 8,
            angle = 90,
            position = position_dodge(width = 0.9), 
            vjust = 1, 
            color = "white") +
  
    facet_wrap(~Application_Type, scale = "fixed")


sex_age_fundability
# Save the plot
ggsave(filename = paste0("C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note05/ethnicity_age_fundability.png"), 
       plot = sex_age_fundability, width = 14, height = 12, dpi = 600, bg = "white")
```

