---
title: "Bias in EPSRC Peer Review - HESA Comparison Charts"
author: 'Contributors: Ged Hiscoke & Dakota Langhals, Royal Statistical Society'
date: "Version: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    code_folding: hide
    toc_float: yes
  pdf_document: default
  word_document:
    reference_docx: "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/r_style_template.docx"
    highlight: null
subtitle: 'Status: Final'
---

```{r setup, include=FALSE}

##Universal Document settings
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12)


## Libraries
library(dplyr)      # For tidy coding
library(kableExtra) # For tables
library(ggplot2)    # For plotting
library(patchwork)  # For plot displays
library(tidyr)      # For tidy coding
library(ggforce)    # For tidy labelling on plots 
library(gridExtra)  # For display of joint plots
library(knitr)
library(openxlsx)   # for saving tables as excel files
library(scales)     # For commas in plot labels
library(binom)
library(purrr)
library(boot)
library(dplyr)
library(reshape2)
library(knitr)
library(vcd)






## Set key file locations
## Directory for saving local high resolution copies of plots etc
data_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Datasets/"
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/HESA Comparison"



## Set key file names
WorkingEnvironment  <- "PROJ2023.002 Environment - WORKING.RData"


## Load needed data
load(file = file.path(data_folder_path, WorkingEnvironment))



## Define the RSS colour palette

RRSSgold      <- "#D3A41B" 
RSSblue       <- "#004B6D" 
RSSred        <- "#cc0000" 
RSStaupe      <- "#a6a086" 
RSSgreen      <- "#009933" 
RSSpurple     <- "#7879b9" 
RSSdark       <- "#393f4d" 
RSSlightblue    <- "#6db4c0" 
RSSseablue       <- "#287fc3" 
RSSorange     <- "#d86027"
RSSlightgrey  <- "#bdbdbd"
RSSmidgrey    <- "#939393"
RSSdarkkgrey  <- "#7e7e7e"
RSSjungle     <- "#1f3d0c"
RSSmint       <- "#5C9090" 
RSSherb       <- "#689E88"
RSSbright     <- "#8D31A4"
RSSlime       <- "#7DCF7A"
RSSviolet     <- "#614182"


##Variable specific colours
col_male      <- "#004b6d"
col_female    <- "#a6700a"
col_uk        <- "#393f4d"
col_nonuk     <- "#614182"
col_dis       <- "#005b49"
col_nodis     <- "#8e8e00"
col_white     <- "#7f3c21"
col_ethmin    <- "#3a3877"
col_unknown   <- "#bdbdbd"
col_notdisclosed  <- "#640106"
col_age1 <- "#b6d6b3"
col_age2 <- "#87bd00"
col_age3 <- "#299907"
col_age4 <- "#146f03"
col_age5 <- "#054907"
col_age6 <- "#0f2808"
col_eth1 <- "#a7a7c1"
col_eth2 <- "#74749d"
col_eth3 <- "#1e3399"
col_eth4 <- "#090771"
col_eth5 <- "#37163e"
col_eth6 <- "#59095f"
col_eth7 <- "#34086e"
```

# {.tabset}


## Defining Functions

```{r}
calculate_HESA_proportions.f <- function(df, group_col) {
  # Function to calculate the confidence interval and handle zero values
  
  df2 <- df %>% filter(Year != "2012/2013" & Year != "2013/2014")
  calculate_ci <- function(fte_sum, total_fte) {
    if (fte_sum == 0) {
      return(c(0, 0))
    } else {
      ci <- binom.confint(fte_sum, total_fte, conf.level = 0.95, methods = "wilson")
      return(c(ci$lower, ci$upper))
    }
  }
  
  # Filter the dataset for Role == "Teach and Research" - this is the most directly comparable to the EPSRC data, as they are the most likely to apply for funding with EPSRC.
  df_teach_and_research <- df2 %>%
    filter(Role == "Teach and Research")
  
  # Group by Year and the specified column and sum the FTE for "Teach and Research" role
  grouped_df_teach_and_research <- df_teach_and_research %>%
    group_by(Year, !!sym(group_col)) %>%
    summarise(FTE_sum = sum(FTE), .groups = 'drop')
  
  # Calculate the total FTE per Year for "Teach and Research" role
  total_fte_per_year_teach_and_research <- grouped_df_teach_and_research %>%
    group_by(Year) %>%
    summarise(Total_FTE = sum(FTE_sum), .groups = 'drop')
  
  # Merge the total FTE back with the grouped data for "Teach and Research" role
  proportion_df_teach_and_research <- grouped_df_teach_and_research %>%
    left_join(total_fte_per_year_teach_and_research, by = "Year") %>%
    mutate(PropTR = FTE_sum / Total_FTE) %>%
    rowwise() %>%
    mutate(
      CI = list(calculate_ci(FTE_sum, Total_FTE)),
      CI_L_TR = CI[1],
      CI_U_TR = CI[2]
    ) %>%
    ungroup() %>%
    select(-CI)  # Remove the CI column containing list elements
  
  return(proportion_df_teach_and_research)
}


```

```{r confidence_interval_function}
##Function-----------------------------------------
# Function to calculate bootstrap CI for proportions
boot_ci <- function(data, indices) {
  sampled_data <- data[indices]
  prop <- mean(sampled_data)
  return(prop)
}

# helper Function to compute bootstrapped confidence intervals
compute_ci <- function(data) {
  if (length(data) < 2) {
    return(data.frame(CI_Lower = NA, CI_Upper = NA))
  }
  
  boot_result <- boot(data, boot_ci, R = 1000)
  ci <- boot.ci(boot_result, type = "basic")
  return(data.frame(CI_Lower = ci$basic[4], CI_Upper = ci$basic[5]))
}

# Generalized function to compute counts, proportions, and CIs by group and demographic
compute_statistics <- function(df, group_col, demo_col) {
  df <- df %>%
    rename(Group = !!sym(group_col), Demographic = !!sym(demo_col))
  
  groups <- unique(df$Group)
  results_list <- list()
  
  for (group in groups) {
    group_df <- df %>% filter(Group == group)
    
    result <- group_df %>%
      group_by(Demographic) %>%
      summarise(
        Count = n(),
        Proportion = Count / nrow(group_df)
      ) %>%
      ungroup() %>%
      rowwise() %>%
      mutate(
        CI = list(compute_ci(group_df$Demographic == Demographic)),
        CI_L = CI$CI_Lower,
        CI_U = CI$CI_Upper
      ) %>%
      select(-CI) %>%
      ungroup() %>%
      mutate(Year = group)
    
    results_list[[group]] <- result
  }
  
  results_df <- bind_rows(results_list) %>%
    mutate(Group = demo_col)
  
  results_df <- results_df %>%
    rename(!!sym(demo_col) := Demographic)
  
  return(results_df)
}

```


```{r}
plot_HESA_Comparison.f <- function(HESA_df, EPSRC_df, demographic, c1, c2, ribbon_fade = 0.5, title_text) {
  
  # Define colors
  colors <- c(
    "HESA: Teaching and Research (95% Confidence Interval)" = c1,
    "EPSRC" = c2
  )
  
  # Prepare the HESA dataframe
  HESA_prepared <- HESA_df %>% 
    filter(!!sym(demographic) != "Unknown" & !!sym(demographic) != "Other")
  
  # Prepare the EPSRC dataframe
  EPSRC_prepared <- EPSRC_df %>% 
    filter(!!sym(demographic) != "Unknown" & !!sym(demographic) != "Not disclosed")
  
  # Plotting
  ggplot() +
    # Ribbon for HESA data
    geom_ribbon(
      data = HESA_prepared,
      aes(
        x = Year,
        ymin = CI_L_TR,
        ymax = CI_U_TR,
        fill = "HESA: Teaching and Research (95% Confidence Interval)",
        group = 1
      ),
      alpha = ribbon_fade
    ) +
    # Points for EPSRC data
    geom_point(
      data = EPSRC_prepared,
      aes(
        x = Year,
        y = Proportion,
        color = "EPSRC"
      ),
      shape = 19,
      size = 4
    ) +
    # Error bars for EPSRC data
    geom_errorbar(
      data = EPSRC_prepared,
      aes(
        x = Year,
        ymin = CI_L,
        ymax = CI_U,
        color = "EPSRC"
      ),
      width = 0.2,
      size = 1
    ) +
    # Facet by demographic
    facet_wrap(as.formula(paste("~", demographic)), scales = "free_y", ncol = 2) +
    # Y-axis labels as percentages
    scale_y_continuous(limits = c(0, 0.45), labels = scales::percent_format(accuracy = 1)) +
    # Custom color scale for EPSRC
    scale_color_manual(values = colors, name = NULL) +
    # Custom fill scale for HESA
    scale_fill_manual(
      values = c("HESA: Teaching and Research (95% Confidence Interval)" = c1),
      name = NULL
    ) +
    # Labels and themes
    labs(
      title = title_text,
      y = "Proportions",
      x = "Year"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"), # Center title, add space below
      axis.title.x = element_text(size = 22, margin = margin(t = 15), face = "bold"), # Add space above x-axis title
      axis.title.y = element_text(size = 22, margin = margin(r = 15), face = "bold"), # Add space to the right of y-axis title
      plot.margin = margin(t = 20, r = 10, b = 10, l = 10), # Add extra space around the plot
      axis.text.x = element_text(size = 16, face = "bold", angle = 90), # Make axis easier to read
      axis.text.y = element_text(size = 16, face = "bold"),  # Make axis easier to read
      strip.text = element_text(size = 16, face = "italic"),
      legend.position = "top",
      legend.text = element_text(size = 16, face = "bold")
    )
}

```

## Sex

<br>

#### Comparing EPSRC Applicants with HESA Population, by Sex

(2014/15 to 2022/23. All Unique Researchers)

```{r}
hesa.df <- calculate_HESA_proportions.f(HESA$Sex.df %>% filter(Sex != "Other", Sex != "Unknown"), "Sex")

hesa.df$Year <- substr(hesa.df$Year, 1, 4)

df <- EPSRC$Investigators.df %>% filter(Sex != "Not Disclosed", Sex != "Unknown") %>% 
  group_by(FYDecisionDate, AnonINV) %>%
  distinct(AnonINV, .keep_all = TRUE) %>%    # Keep only one unique AnonINV per year
  ungroup()

epsrc.df <- compute_statistics(df, "FYDecisionDate", "Sex")

epsrc.df$Year <- substr(epsrc.df$Year, 1, 4)



HESA_sex <- plot_HESA_Comparison.f(
  HESA_df = hesa.df %>% filter(Sex == "Female"),
  EPSRC_df = epsrc.df %>% filter(Sex == "Female"),
  demographic = "Sex",
  RSSgold,
  "black",
  title_text = "Proportion of EPSRC Applicants Who Are Female\nCompared to HESA Population"
)

print(HESA_sex)


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/HESA Comparison"

ggsave(filename = paste0(figure_folder_path, "/sex_HESA_comparison.png"), 
       plot = HESA_sex, width = 12, height = 8, dpi = 800, bg = "white")
```


## Ethnicity

#### Comparing EPSRC Applicants with HESA Population, by Binary Ethnicity

(2014/15 to 2022/23. All Unique Researchers)

```{r}
hesa.df <- calculate_HESA_proportions.f(HESA$Ethnicity.df %>% filter(EthnicityBinary != "Other", EthnicityBinary != "Unknown"), "EthnicityBinary")

hesa.df$Year <- substr(hesa.df$Year, 1, 4)

df <- EPSRC$Investigators.df %>% filter(EthnicityBinary != "Not Disclosed", EthnicityBinary != "Unknown") %>% 
  group_by(FYDecisionDate, AnonINV) %>%
  distinct(AnonINV, .keep_all = TRUE) %>%    # Keep only one unique AnonINV per year
  ungroup()

epsrc.df <- compute_statistics(df, "FYDecisionDate", "EthnicityBinary")

epsrc.df$Year <- substr(epsrc.df$Year, 1, 4)



HESA_ethnicity <- plot_HESA_Comparison.f(
  HESA_df = hesa.df %>% filter(EthnicityBinary == "Ethnic Minority"),
  EPSRC_df = epsrc.df %>% filter(EthnicityBinary == "Ethnic Minority"),
  demographic = "EthnicityBinary",
  RSSgold,
  "black",
  title_text = "Proportion of EPSRC Applicants Who Are Ethnic Minorities\nCompared to HESA Population"
)

print(HESA_ethnicity)

figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/HESA Comparison"

ggsave(filename = paste0(figure_folder_path, "/ethnicity_HESA_comparison.png"), 
       plot = HESA_ethnicity, width = 12, height = 8, dpi = 800, bg = "white")

```

## Age

#### Comparing the 2014/2015 Age Distribution of Researchers in the EPSRC Applicant and HESA Population

(2014/15 to 2022/23. All Unique Researchers)

```{r}
hesa.df <- HESA$Age.df %>% filter(AgeRange != "Other", AgeRange != "Unknown")

# Collapsing age
hesa.df <- hesa.df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35_and_Under", 
                                       AgeRange == "26-35" ~ "35_and_Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56_and_Over",
                                       AgeRange == "66+" ~  "56_and_Over"))

hesa.df <- calculate_HESA_proportions.f(hesa.df, "AgeRangeCollapsed")

hesa.df$Year <- substr(hesa.df$Year, 1, 4)

df <- EPSRC$Investigators.df %>% filter(AgeRange != "Not Disclosed", AgeRange != "Unknown") %>% 
  group_by(FYDecisionDate, AnonINV) %>%
  distinct(AnonINV, .keep_all = TRUE) %>%    # Keep only one unique AnonINV per year
  ungroup()

# Collapsing age
df <- df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35_and_Under", 
                                       AgeRange == "26-35" ~ "35_and_Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56_and_Over",
                                       AgeRange == "66+" ~  "56_and_Over"))

epsrc.df <- compute_statistics(df, "FYDecisionDate", "AgeRangeCollapsed")

epsrc.df$Year <- substr(epsrc.df$Year, 1, 4)



age_colors <- c(
  "35_and_Under" = RSSblue,
  "36-55" = RSSgold,
  "56_and_Over" = RSSred
)

HESA_age <- ggplot() +
  # Ribbon for HESA data
  geom_ribbon(
    data = hesa.df,
    aes(
      x = Year,
      ymin = CI_L_TR,
      ymax = CI_U_TR,
      fill = AgeRangeCollapsed,
      group = AgeRangeCollapsed
    ),
    alpha = 0.3
  ) +
  # Points for EPSRC data
  geom_point(
    data = epsrc.df,
    aes(
      x = Year,
      y = Proportion,
      color = AgeRangeCollapsed
    ),
    size = 3
  ) +
  # Error bars for EPSRC data
  geom_errorbar(
    data = epsrc.df,
    aes(
      x = Year,
      ymin = CI_L,
      ymax = CI_U,
      color = AgeRangeCollapsed
    ),
    width = 0.2,
    size = 1
  ) +
  # Y-axis labels as percentages
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  # Custom scales
  scale_color_manual(
    values = age_colors,
    name = "Age Range (EPSRC)"
  ) +
  scale_fill_manual(
    values = age_colors,
    name = "Age Range (HESA)"
  ) +
  scale_x_discrete(breaks = seq(min(epsrc.df$Year), max(epsrc.df$Year), by = 1))  +
  guides(
    shape = "none",  # Remove shape legend
    fill = guide_legend(order = 1),  # Maintain order for Age Range (HESA)
    color = guide_legend(order = 2)  # Maintain order for Age Range (EPSRC)
  ) +
  # Labels and themes
  labs(
    title = "Proportion of EPSRC Applicants by Age Range\nCompared to HESA Population",
    y = "Proportions",
    x = "Year"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    axis.title.x = element_text(size = 22, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 22, margin = margin(r = 15), face = "bold"),
    axis.text.x = element_text(size = 16, face = "bold", angle = 90),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 16, face = "bold"),
    legend.box = "vertical",          # Arrange legend items horizontally
    legend.box.just = "center",         # Center the legend
    legend.spacing.y = unit(0.3, "cm"), # Adjust vertical spacing between legend elements
    legend.key.height = unit(0.5, "cm") # Adjust height of legend keys for better spacing
   )

print(HESA_age)


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/HESA Comparison"

ggsave(filename = paste0(figure_folder_path, "/age_HESA_comparison.png"), 
       plot = HESA_age, width = 12, height = 8, dpi = 800, bg = "white")

```



## Nationality

#### Comparing EPSRC Applicants with HESA Population, by Nationality

(2014/15 to 2022/23. All Unique Researchers)

```{r}
hesa.df <- calculate_HESA_proportions.f(HESA$Nationality.df %>% filter(NationalityBinary != "Other", NationalityBinary != "Unknown"), "NationalityBinary")

hesa.df$Year <- substr(hesa.df$Year, 1, 4)

df <- EPSRC$Investigators.df %>% filter(NationalityBinary != "Not Disclosed", NationalityBinary != "Unknown") %>% 
  group_by(FYDecisionDate, AnonINV) %>%
  distinct(AnonINV, .keep_all = TRUE) %>%    # Keep only one unique AnonINV per year
  ungroup()

epsrc.df <- compute_statistics(df, "FYDecisionDate", "NationalityBinary")

epsrc.df$Year <- substr(epsrc.df$Year, 1, 4)



HESA_nationality <- plot_HESA_Comparison.f(
  HESA_df = hesa.df %>% filter(NationalityBinary == "Non UK"),
  EPSRC_df = epsrc.df %>% filter(NationalityBinary == "Non UK"),
  demographic = "NationalityBinary",
  RSSgold,
  "black",
  title_text = "Proportion EPSRC Applicants Who Are Not from the UK\nCompared to HESA Population"
)

print(HESA_nationality)

figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/HESA Comparison"

ggsave(filename = paste0(figure_folder_path, "/nationality_HESA_comparison.png"), 
       plot = HESA_nationality, width = 12, height = 8, dpi = 800, bg = "white")
```

## Disability

#### Comparing EPSRC Applicants with HESA Population, by Disclosed Disability

(2014/15 to 2022/23. All Unique Researchers)

```{r}
hesa.df <- calculate_HESA_proportions.f(HESA$Disability.df %>% filter(Disability != "Other", Disability != "Unknown"), "Disability")

hesa.df$Year <- substr(hesa.df$Year, 1, 4)

df <- EPSRC$Investigators.df %>% filter(Disability != "Not Disclosed", Disability != "Unknown") %>% 
  group_by(FYDecisionDate, AnonINV) %>%
  distinct(AnonINV, .keep_all = TRUE) %>%    # Keep only one unique AnonINV per year
  ungroup()

epsrc.df <- compute_statistics(df, "FYDecisionDate", "Disability")

epsrc.df$Year <- substr(epsrc.df$Year, 1, 4)



HESA_disability <- plot_HESA_Comparison.f(
  HESA_df = hesa.df %>% filter(Disability == "Known Disability"),
  EPSRC_df = epsrc.df %>% filter(Disability == "Known Disability"),
  demographic = "Disability",
  RSSgold,
  "black",
  title_text = "Proportion EPSRC Applicants with Known Disabilities\nCompared to HESA Population"
)

print(HESA_disability)

figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/HESA Comparison"

ggsave(filename = paste0(figure_folder_path, "/disability_HESA_comparison.png"), 
       plot = HESA_disability, width = 12, height = 8, dpi = 800, bg = "white")
```


