---
title: "Bias in EPSRC Peer Review - Research Question 8"
author: 'Contributors: Ged Hiscoke & Dakota Langhals, Royal Statistical Society'
date: "Version: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    code_folding: hide
    toc_float: yes
  pdf_document: default
  word_document:
    reference_docx: "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/r_style_template.docx"
    highlight: null
subtitle: 'Status: Final'
---

```{r setup, include=FALSE}

##Universal Document settings
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12, warning = FALSE, message = FALSE)


## Libraries
library(dplyr)      # For tidy coding
library(kableExtra) # For tables
library(ggplot2)    # For plotting
library(patchwork)  # For plot displays
library(tidyr)      # For tidy coding
library(ggforce)    # For tidy labelling on plots 
library(gridExtra)  # For display of joint plots
library(knitr)
library(openxlsx)   # for saving tables as excel files
library(scales)     # For commas in plot labels
library(boot)
library(formatR)
library(purrr)
library(ggrepel)    # For helping with labels on graphs
library(cowplot)
library(forcats)






## Set key file locations
## Directory for saving local high resolution copies of plots etc
data_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Datasets/"
note_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Notes/"
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note08"
table_folder_path   <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Tables/Note08"



## Set key file names
raw_data_file       <- "PROJ2023.002-DATA001.xlsx"
BaselineEnvironment <- "PROJ2023.002 Environment - BASELINE.RData"
CleanedEnvironment  <- "PROJ2023.002 Environment - CLEANED.RData"
Sample20Environment <- "PROJ2023.002 Environment - SAMPLE20.RData"
Sample80Environment <- "PROJ2023.002 Environment - SAMPLE80.RData"
cleanedExcelFile    <- "PROJ2023.002-DATA002.xlsx"
WorkingEnvironment  <- "PROJ2023.002 Environment - WORKING.RData"
UKLFdata            <- "PROJ2023.002-DATA004.xlsx"


## Load needed data
load(file = file.path(data_folder_path, WorkingEnvironment))



## Define the RSS colour palette

RRSSgold      <- "#D3A41B" 
RSSblue       <- "#004B6D" 
RSSred        <- "#cc0000" 
RSStaupe      <- "#a6a086" 
RSSgreen      <- "#009933" 
RSSpurple     <- "#7879b9" 
RSSdark       <- "#393f4d" 
RSSlightblue    <- "#6db4c0" 
RSSseablue       <- "#287fc3" 
RSSorange     <- "#d86027"
RSSlightgrey  <- "#bdbdbd"
RSSmidgrey    <- "#939393"
RSSdarkkgrey  <- "#7e7e7e"
RSSjungle     <- "#1f3d0c"
RSSmint       <- "#5C9090" 
RSSherb       <- "#689E88"
RSSbright     <- "#8D31A4"
RSSlime       <- "#7DCF7A"
RSSviolet     <- "#614182"


##Variable specific colours
col_male      <- "#004b6d"
col_female    <- "#a6700a"
col_uk        <- "#393f4d"
col_nonuk     <- "#614182"
col_dis       <- "#005b49"
col_nodis     <- "#8e8e00"
col_white     <- "#7f3c21"
col_ethmin    <- "#3a3877"
col_unknown   <- "#bdbdbd"
col_notdisclosed  <- "#640106"
col_age1 <- "#b6d6b3"
col_age2 <- "#87bd00"
col_age3 <- "#299907"
col_age4 <- "#146f03"
col_age5 <- "#054907"
col_age6 <- "#0f2808"
col_eth1 <- "#a7a7c1"
col_eth2 <- "#74749d"
col_eth3 <- "#1e3399"
col_eth4 <- "#090771"
col_eth5 <- "#37163e"
col_eth6 <- "#59095f"
col_eth7 <- "#34086e"
```

# {.tabset}

## Data preparation

```{r datacleaning}
# Identify the list of Resubmitters ---------------------------------------

df <- EPSRC$Investigators.df
# Load necessary library
library(dplyr)

# Sample data frame
# df <- data.frame(AnonINV = ..., FYDecisionDate = ..., GrantOutcome = ..., FYReceivedDate = ...)

# Function to convert academic year to a start year
convert_academic_year <- function(academic_year) {
  as.numeric(substr(academic_year, 1, 4))
}

# Convert FYDecisionDate and FYReceivedDate to start years
df <- df %>%
  mutate(DecisionYear = sapply(FYDecisionDate, convert_academic_year),
         SubmissionYear = sapply(FYReceivedDate, convert_academic_year))

# Initialize an empty data frame to store results
results <- data.frame(AnonINV = character(),
                      OriginalDecisionYear = numeric(),
                      NextSubmissionYear = numeric(),
                      YearsToNextSubmission = numeric(),
                      stringsAsFactors = FALSE)

# Iterate over each unique AnonINV
unique_invs <- unique(df$AnonINV)

for (inv in unique_invs) {
  # Get the applications of this investigator, sorted by SubmissionYear
  inv_applications <- df %>%
    filter(AnonINV == inv) %>%
    arrange(SubmissionYear)
  
  # Check if the investigator has any funded applications before the first unfunded application
  has_prior_success <- any(inv_applications$GrantOutcome[inv_applications$SubmissionYear < min(inv_applications$SubmissionYear[inv_applications$GrantOutcome != 'Funded'])] == 'Funded')
  
  if (!has_prior_success) {
    # Track the last unfunded decision year
    last_unfunded_year <- NA
    last_unfunded_agerange <- NA
    
    for (i in 1:nrow(inv_applications)) {
      row <- inv_applications[i, ]
      
      if (row$GrantOutcome != 'Funded') {
        # Update the last unfunded decision year
        last_unfunded_year <- row$DecisionYear
        last_unfunded_agerange <- row$AgeRange
        
      } else if (!is.na(last_unfunded_year) && row$SubmissionYear > last_unfunded_year) {
        # Calculate the number of years to the next submission after an unfunded decision
        years_to_next_submission <- row$SubmissionYear - last_unfunded_year
        
        # Store the result
        results <- rbind(results, data.frame(
          AnonINV = inv,
          OriginalDecisionYear = last_unfunded_year,
          NextSubmissionYear = row$SubmissionYear,
          YearsToNextSubmission = years_to_next_submission,
          AgeRange = last_unfunded_agerange,
          stringsAsFactors = FALSE
        ))
        
        # Reset the last unfunded year to avoid counting the same unfunded decision multiple times
        last_unfunded_year <- NA
        last_unfunded_agerange <- NA
      }
    }
  }
}
resubmitters <- results

# Identify the list of non-resubmitters-----------------------------------
# Function to convert academic year to a start year
convert_academic_year <- function(academic_year) {
  as.numeric(substr(academic_year, 1, 4))
}

# Convert FYDecisionDate and FYReceivedDate to start years
df <- df %>%
  mutate(DecisionYear = sapply(FYDecisionDate, convert_academic_year),
         SubmissionYear = sapply(FYReceivedDate, convert_academic_year))

# Filter for applicants who made only one submission and it was not funded
no_resubmissions <- df %>%
  group_by(AnonINV) %>%
  filter(n() == 1 & GrantOutcome != 'Funded') %>%
  ungroup()

# Add in demographic details ----------------------------------------------
Q8data.df <- no_resubmissions %>% select(AnonINV, AgeRange)
Q8data.df$Resubmission <- "No Resubmission"
resubmitters$Resubmission <- "Resubmission"
resubmitters <- resubmitters %>% select(AnonINV, AgeRange, Resubmission)
Q8data.df <- Q8data.df %>% rbind(resubmitters) 
uniquedf <- EPSRC$Investigators.df %>% select(Sex, EthnicityBinary, NationalityBinary, EthnicGroup, Disability, AnonINV)
uniquedf <- distinct(uniquedf)
Q8data.df <- Q8data.df %>%left_join(uniquedf , AnonINV, by = "AnonINV")

# Collapsing age to align with ATI and make interactions more easily interpretable
Q8data.df <- Q8data.df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35 and Under", 
                                       AgeRange == "26-35" ~ "35 and Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56 and Over",
                                       AgeRange == "66+" ~  "56 and Over",
                                       AgeRange == "Unknown" ~ "Unknown"))

# Intersectional demographics
Q8data.df$Sex_Ethnicity <- interaction(Q8data.df$Sex, Q8data.df$EthnicityBinary)
Q8data.df$Sex_Age <- interaction(Q8data.df$Sex, Q8data.df$AgeRangeCollapsed)
Q8data.df$Ethnicity_Age <- interaction(Q8data.df$EthnicityBinary, Q8data.df$AgeRangeCollapsed)




```


## Proportions of Resubmitting and Non-resubmitting Populations Belonging to Demographic Groups

```{r}
##Function-----------------------------------------

# Generalized function to compute counts, proportions, and CIs by group and demographic
compute_statistics <- function(df, group_col, demo_col, driver = "Uncertain") {
  
  # Function to calculate bootstrap CI for proportions
  boot_ci <- function(data, indices) {
    sampled_data <- data[indices]
    prop <- mean(sampled_data)
    return(prop)
  }

  # Helper function to compute bootstrapped confidence intervals
  compute_ci <- function(data) {
    if (length(data) < 2 || all(is.na(data))) {
      # If there are fewer than 2 observations or data is NA, return NA for CI
      return(data.frame(CI_Lower = NA, CI_Upper = NA))
    }
    
    boot_result <- boot(data, boot_ci, R = 1000)
    ci <- boot.ci(boot_result, type = "basic")
    
    if (is.null(ci)) {
      return(data.frame(CI_Lower = NA, CI_Upper = NA))  # Handle case where bootstrapping fails
    }
    
    return(data.frame(CI_Lower = ci$basic[4], CI_Upper = ci$basic[5]))
  }

  df <- df %>%
    rename(Group = !!sym(group_col), Demographic = !!sym(demo_col))
  
  groups <- unique(df$Group)
  results_list <- list()
  
  for (group in groups) {
    group_df <- df %>% filter(Group == group)
    
    result <- group_df %>%
      group_by(Demographic) %>%
      summarise(
        Count = n(),
        Proportion = Count / nrow(group_df)
      ) %>%
      ungroup() %>%
      rowwise() %>%
      mutate(
        CI = list(compute_ci(group_df$Demographic == Demographic)),
        CI_L = CI$CI_Lower,
        CI_U = CI$CI_Upper
      ) %>%
      select(-CI) %>%
      ungroup() %>%
      mutate(Year = group)
    
    results_list[[group]] <- result
  }
  
  results_df <- bind_rows(results_list) %>%
    mutate(Group = demo_col)
  
  results_df$Driver <- driver
  return(results_df)
}



```

```{r}
# Function to round columns as specified
round_columns <- function(df) {
  df <- df %>%
    mutate(
      Count = ifelse(Count < 6, "\u22645", round(Count / 5) * 5),
      across(where(is.numeric) & !c(Count), ~ ifelse(Count == "\u22645", "~", round(., 3)))
    )
  return(df)
}

```


#### Table: Proportion of Resubmitting Population by Levels of Demographics

```{r}

df_sex <- Q8data.df %>% filter(Sex != "Unknown" & Sex != "Not Disclosed")
df_ethnicity <- Q8data.df %>% filter(EthnicityBinary != "Unknown" & EthnicityBinary != "Not Disclosed")
df_nationality <- Q8data.df %>% filter(NationalityBinary != "Unknown")
df_disability <- Q8data.df %>% filter(Disability != "Unknown" & Disability != "Not Disclosed")
df_age <- Q8data.df %>% filter(AgeRangeCollapsed != "Unknown" & AgeRangeCollapsed != "Not Disclosed", !is.na(AgeRangeCollapsed))


t1 <-compute_statistics(df_sex, "Resubmission", "Sex","")
t1<- rbind(t1,compute_statistics(df_ethnicity, "Resubmission", "EthnicityBinary",""))
t1<- rbind(t1,compute_statistics(df_nationality, "Resubmission", "NationalityBinary",""))
t1<- rbind(t1,compute_statistics(df_disability, "Resubmission", "Disability",""))
t1<- rbind(t1,compute_statistics(df_age, "Resubmission", "AgeRangeCollapsed",""))


table8.01 <- t1 %>% 
  select(-Driver) %>% 
  round_columns()
  
kable(table8.01)
```

#### Plot: Proportion of Resubmitting Population by Levels of Demographics

```{r, fig.height = 12, fig.width = 12}
theme_my_defaults <- function() {
  theme(
    text = element_text(family = "Frutiger LT Std"),
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"), # Center title, add space below
    axis.title.x = element_text(size = 22, margin = margin(t = 15), face = "bold"), # Add space above x-axis title
    axis.title.y = element_text(size = 22, margin = margin(r = 15), face = "bold"), # Add space to the right of y-axis title
    # plot.margin = margin(t = 20, r = 10, b = 10, l = 10), # Add extra space around the plot
    axis.text.x = element_text(size = 16, face = "bold"), # Make axis easier to read
    axis.text.y = element_text(size = 16, face = "bold"),  # Make axis easier to read
    plot.title.position = "plot",
    panel.background = element_rect(fill = "white"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18, face = "bold")
  )
}


plot_colors <-
  c(
    "Male"= RSSred,      
    "Female"= RSSblue,    
    "UK"=col_uk,        
    "Non UK"=col_nonuk,   
    "Undisclosed Nationality" = RSSmidgrey,
    "Known Disability"=col_dis   ,    
    "No Known Disability"=col_nodis  ,   
    "White"=col_white  ,   
    "Ethnic Minority"=col_ethmin,
    "56 and Over" = col_5665,
    "36-55" = col_3645,
    "35 and Under" = col_2635
    
  )

df <- t1 %>% 
  select(-Driver) %>% 
  rename(Resubmission_Status = Year) 

df <- df %>% mutate(
  Demographic = case_when(
  Demographic == "Not Disclosed" ~ "Undisclosed Nationality",
  TRUE ~ Demographic
))


# Convert Group to a factor with a custom order
df <- df %>%
  mutate(Group = factor(Group, levels = c("AgeRangeCollapsed", "Disability", "EthnicityBinary", "NationalityBinary", "Sex")))

# Reorder Demographic based on Group
df <- df %>%
  mutate(Demographic = fct_reorder(Demographic, as.numeric(Group)))



### Calculate the baseline proportions in each group that receive at least one rejection, then add to df for plotting
temp <- df %>% 
  group_by(Group) %>% 
  mutate(
    total_group = sum(Count)
  ) %>% 
  ungroup() %>% 
  group_by(Demographic) %>% 
  mutate(
    total_demographic = sum(Count)
  ) %>% 
  ungroup() %>% 
  mutate(
    Proportion = total_demographic / total_group,
    Count = total_demographic,
    Resubmission_Status = "(Baseline)",
    CI_L = NA,
    CI_U = NA
  ) %>% 
  ungroup() %>% 
  select(-c(total_group, total_demographic))

temp <- temp %>% distinct(Group, Demographic, .keep_all = TRUE)

df <- rbind(df, temp)



# Plot the CIs, faceted by Group, colored by Demographic, with dodged positions

fig8.01 <- ggplot(df, aes(x = Demographic, y = Proportion, color = Demographic, shape = Resubmission_Status)) +
  geom_point(position = position_dodge(width = 1), size = 4) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
      position = position_dodge(width = 1), 
      width = 0.5,
      size = 1.2,
      alpha = 0.5) +
  # geom_text_repel(show.legend = FALSE,
  #                 aes(label = scales::percent(Proportion, accuracy = 1)),
  #                 position = position_dodge(width = 1), 
  #                 vjust = -1, 
  #                 size = 5,
  #                 force = 2, # Adjust the force to control label repelling
  #                 max.overlaps = Inf, # Prevents any labels from being removed
  #                 box.padding = 0.7, # Adds padding around labels
  #                 point.padding = 0.2) +
  scale_color_manual(values = plot_colors) +
  labs(title = "Proportions of Resubmitting and Non-resubmitting Populations\nby Demographic Group",
       x = "Demographic",
       y = "Proportion of Resubmitting or Non-resubmitting Population") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip() + 
  guides(color = guide_legend(reverse = TRUE),
         shape = guide_legend(reverse = TRUE)) +
  theme_my_defaults()


fig8.01

figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note08"

ggsave(filename = paste0(figure_folder_path, "/dem_proportions_within_resub_pop.png"), 
       plot = fig8.01, width = 12, height = 12, dpi = 800, bg = "white")

```

#### Table: Proportion of Resubmitting Population by Levels of Intersectional Demographics

```{r}
# Intersectional demographics
## Removing unknown and not disclosed values
df <- Q8data.df %>%
  filter(!Sex == "Unknown" &
           !EthnicityBinary == "Unknown" &
           !NationalityBinary == "Unknown" &
           !AgeRange == "Unknown" &
           !Disability == "Unknown" &
           !Sex == "Not Disclosed" &
           !EthnicityBinary == "Not Disclosed" &
           !AgeRange == "Not Disclosed" &
           !Disability == "Not Disclosed") %>% 
  droplevels()


df$Sex_Ethnicity <- interaction(df$Sex, df$EthnicityBinary)
df$Sex_Age <- interaction(df$Sex, df$AgeRangeCollapsed)
df$Ethnicity_Age <- interaction(df$EthnicityBinary, df$AgeRangeCollapsed)

df$Sex_Ethnicity_Age <- interaction(df$Sex, df$EthnicityBinary, df$AgeRangeCollapsed)

t1 <- compute_statistics(df, "Resubmission", "Sex_Ethnicity","")
t2 <- compute_statistics(df, "Resubmission", "Sex_Age","")
t3 <- compute_statistics(df, "Resubmission", "Ethnicity_Age","")

t4 <- compute_statistics(df, "Resubmission", "Sex_Ethnicity_Age", "")


intersections_table1 <- t1 %>% 
  select(-Driver) %>% 
  round_columns()
  
kable(intersections_table1)

intersections_table2 <- t2 %>% 
  select(-Driver) %>% 
  round_columns()
  
kable(intersections_table2)

intersections_table3 <- t3 %>% 
  select(-Driver) %>% 
  round_columns()

kable(intersections_table3)

intersections_table4 <- t4 %>% 
  select(-Driver) %>% 
  round_columns()
  
kable(intersections_table4)
```

#### Plot: Proportion of Resubmitting Population by Levels of Intersectional Demographics

```{r, fig.height = 12, fig.width = 12}
# Define custom colors
plot_colors <-
  c(
    "Male.White"= RSSblue,      
    "Female.White"= RSSred,
    "Male.Ethnic Minority"=RSSlightblue,
    "Female.Ethnic Minority"=RSSgold,     
    "Female.35_and_Under"=RSSred   ,    
    "Female.36-55"=RSSgold  ,   
    "Female.56_and_Over"=RSSorange  ,   
    "Male.35_and_Under"=RSSblue  ,  
    "Male.36-55"=RSSviolet  , 
    "Male.56 and Over"=RSSlightblue  ,
    "White.35 and Under"=RSSblue ,
    "White.36-55"=RSSviolet ,
    "White.56 and Over"=RSSlightblue ,
    "Ethnic Minority.35 and Under"=RSSred ,
    "Ethnic Minority.36 55"=RSSgold ,
    "Ethnic Minority.56 and Over"=RSSorange
  )

df1 <- t1 %>% 
  select(-Driver) %>% 
  rename(Resubmission_Status = Year) 

df2 <- t2 %>% 
  select(-Driver) %>% 
  rename(Resubmission_Status = Year) 

df3 <- t3 %>% 
  select(-Driver) %>% 
  rename(Resubmission_Status = Year) 

df4 <- t4 %>% 
  select(-Driver) %>% 
  rename(Resubmission_Status = Year) 

### Calculate the baseline proportions in each group that receive at least one rejection, then add to df4 for plotting
temp <- df4 %>% 
  group_by(Group) %>% 
  mutate(
    total_group = sum(Count)
  ) %>% 
  ungroup() %>% 
  group_by(Demographic) %>% 
  mutate(
    total_demographic = sum(Count)
  ) %>% 
  ungroup() %>% 
  mutate(
    Proportion = total_demographic / total_group,
    Count = total_demographic,
    Resubmission_Status = "(Baseline)",
    CI_L = NA,
    CI_U = NA
  ) %>% 
  ungroup() %>% 
  select(-c(total_group, total_demographic))

temp <- temp %>% distinct(Group, Demographic, .keep_all = TRUE)

df4 <- rbind(df4, temp)



# Plot the CIs, faceted by Group, colored by Demographic, with dodged positions

intersections_fig1 <- ggplot(df1, aes(x = Demographic, y = Proportion, color = Demographic, shape = Resubmission_Status)) +
  geom_point(position = position_dodge(width = 1), size = 4) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
      position = position_dodge(width = 1), 
      width = 0.5,
      size = 1.2,
      alpha = 0.5) +
  geom_text_repel(aes(label = scales::percent(Proportion, accuracy = 1)),
                  position = position_dodge(width = 1), 
                  vjust = -1, 
                  size = 5,
                  force = 2, # Adjust the force to control label repelling
                  max.overlaps = Inf, # Prevents any labels from being removed
                  box.padding = 0.7, # Adds padding around labels
                  point.padding = 0.2) +
  scale_color_manual(values = plot_colors) +
  labs(title = "Proportions of Resubmitting and Non-resubmitting Populations\nby Demographic Group",
       x = "Demographic",
       y = "Proportion of Resubmitting or\nNon-resubmitting Population") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip() +
  theme_my_defaults() +
  guides(color = "none")



intersections_fig2 <- ggplot(df2, aes(x = Demographic, y = Proportion, color = Demographic, shape = Resubmission_Status)) +
  geom_point(position = position_dodge(width = 1), size = 4) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
      position = position_dodge(width = 1), 
      width = 0.5,
      size = 1.2,
      alpha = 0.5) +
  geom_text_repel(aes(label = scales::percent(Proportion, accuracy = 1)),
                  position = position_dodge(width = 1), 
                  vjust = -1, 
                  size = 5,
                  force = 2, # Adjust the force to control label repelling
                  max.overlaps = Inf, # Prevents any labels from being removed
                  box.padding = 0.7, # Adds padding around labels
                  point.padding = 0.2) +
  scale_color_manual(values = plot_colors) +
  labs(title = "Proportions of Resubmitting and Non-resubmitting Populations\nby Demographic Group",
       x = "Demographic",
       y = "Proportion of Resubmitting or\nNon-resubmitting Population") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip() +
  theme_my_defaults() +
  guides(color = "none")


intersections_fig3 <- ggplot(df3, aes(x = Demographic, y = Proportion, color = Demographic, shape = Resubmission_Status)) +
  geom_point(position = position_dodge(width = 1), size = 4) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
      position = position_dodge(width = 1), 
      width = 0.5,
      size = 1.2,
      alpha = 0.5) +
  geom_text_repel(aes(label = scales::percent(Proportion, accuracy = 1)),
                  position = position_dodge(width = 1), 
                  vjust = -1, 
                  size = 5,
                  force = 2, # Adjust the force to control label repelling
                  max.overlaps = Inf, # Prevents any labels from being removed
                  box.padding = 0.7, # Adds padding around labels
                  point.padding = 0.2) +
  scale_color_manual(values = plot_colors) +
  labs(title = "Proportions of Resubmitting and Non-resubmitting Populations\nby Demographic Group",
       x = "Demographic",
       y = "Proportion of Resubmitting or\nNon-resubmitting Population") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip() +
  theme_my_defaults() +
  guides(color = "none")




plot_colors <-
  c(
    "Male.White.56 and Over"= RSSblue,      
    "Female.White.56 and Over"= RSSred,
    "Male.Ethnic Minority.56 and Over"=RSSlightblue,
    "Female.Ethnic Minority.56 and Over"=RSSgold, 
    "Male.White.36-55"= RSSviolet,      
    "Female.White.36-55"= RSSorange,
    "Male.Ethnic Minority.36-55"= RSSgreen,
    "Female.Ethnic Minority.36-55"=RSSdark, 
    "Male.White.35 and Under"= RSSpurple,      
    "Female.White.35 and Under"= RSStaupe,
    "Male.Ethnic Minority.35 and Under"=RSSseablue,
    "Female.Ethnic Minority.35 and Under"=RSSmint 
  )

intersections_fig4 <- ggplot(df4, aes(x = Demographic, y = Proportion, color = Demographic, shape = Resubmission_Status)) +
  geom_point(position = position_dodge(width = 1), size = 4) +
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
      position = position_dodge(width = 1), 
      width = 0.5,
      size = 1.2,
      alpha = 0.5) +
  scale_color_manual(values = plot_colors) +
  labs(title = "Proportions of Resubmitting and Non-resubmitting Populations\nby Demographic Group",
       x = "Demographic",
       y = "Proportion of Resubmitting or\nNon-resubmitting Population") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_flip() +
  theme(
    text = element_text(family = "Frutiger LT Std"),
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    axis.title.x = element_text(size = 22, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 22, margin = margin(r = 15), face = "bold"),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    plot.title.position = "plot",
    panel.background = element_rect(fill = "white"),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18, face = "bold"),
    legend.position = "top",
    legend.justification = "center",
    legend.direction = "horizontal",
    legend.box = "vertical",
    legend.box.just = "center"
  ) +
  guides(
    color = "none",
    shape = guide_legend(
      reverse = TRUE,
      title.position = "top",
      label.position = "bottom",
      title.hjust = 0.5,
      label.hjust = 0.5
    )
  )


print(intersections_fig1)
print(intersections_fig2)
print(intersections_fig3)
print(intersections_fig4)



figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/Note08"

ggsave(filename = paste0(figure_folder_path, "/sex+ethnicity_proportions_within_resub_pop.png"),
       plot = intersections_fig1, width = 12, height = 12, dpi = 600, bg = "white")

ggsave(filename = paste0(figure_folder_path, "/sex+age_proportions_within_resub_pop.png"),
       plot = intersections_fig2, width = 12, height = 12, dpi = 600, bg = "white")

ggsave(filename = paste0(figure_folder_path, "/ethnicity+age_proportions_within_resub_pop.png"),
       plot = intersections_fig3, width = 12, height = 12, dpi = 600, bg = "white")

ggsave(filename = paste0(figure_folder_path, "/sex+ethnicity+age_proportions_within_resub_pop.png"),
       plot = intersections_fig4, width = 12, height = 12, dpi = 600, bg = "white")

```






















