---
title: "Bias in EPSRC Peer Review - Research Question 2"
author: 'Contributors: Ged Hiscoke & Dakota Langhals, Royal Statistical Society'
date: "Version: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    code_folding: hide
    toc_float: yes
  pdf_document: default
  word_document:
    reference_docx: "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/r_style_template.docx"
    highlight: null
subtitle: 'Status: Final'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 12, fig.width = 12, warning = FALSE, message = FALSE)

library(dplyr)
library(kableExtra)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(openxlsx)


## Set key file locations
## Directory for saving local high resolution copies of plots etc
data_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Datasets/"
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 02"



## Set key file names
WorkingEnvironment  <- "PROJ2023.002 Environment - WORKING.RData"

load(file = file.path(data_folder_path, WorkingEnvironment))

## Define the RSS colour palette

RRSSgold      <- "#D3A41B" 
RSSblue       <- "#004B6D" 
RSSred        <- "#cc0000" 
RSStaupe      <- "#a6a086" 
RSSgreen      <- "#009933" 
RSSpurple     <- "#7879b9" 
RSSdark       <- "#393f4d" 
RSSlightblue    <- "#6db4c0" 
RSSseablue       <- "#287fc3" 
RSSorange     <- "#d86027"
RSSlightgrey  <- "#bdbdbd"
RSSmidgrey    <- "#939393"
RSSdarkkgrey  <- "#7e7e7e"
RSSjungle     <- "#1f3d0c"
RSSmint       <- "#5C9090" 
RSSherb       <- "#689E88"
RSSbright     <- "#8D31A4"
RSSlime       <- "#7DCF7A"
RSSviolet     <- "#614182"


##Variable specific colours
col_male      <- "#004b6d"
col_female    <- "#a6700a"
col_uk        <- "#393f4d"
col_nonuk     <- "#614182"
col_dis       <- "#005b49"
col_nodis     <- "#8e8e00"
col_white     <- "#7f3c21"
col_ethmin    <- "#3a3877"
col_unknown   <- "#bdbdbd"
col_notdisclosed  <- "#640106"
col_age1 <- "#b6d6b3"
col_age2 <- "#87bd00"
col_age3 <- "#299907"
col_age4 <- "#146f03"
col_age5 <- "#054907"
col_age6 <- "#0f2808"
col_eth1 <- "#a7a7c1"
col_eth2 <- "#74749d"
col_eth3 <- "#1e3399"
col_eth4 <- "#090771"
col_eth5 <- "#37163e"
col_eth6 <- "#59095f"
col_eth7 <- "#34086e"

```

# {.tabset}

## Sex

#### Comparing All EPSRC Population Proportions Across Roles, by Year, by Female Proportion
```{r}

df_panelist <- CI_Tables$Panelist$Sex.df
df_awardee <- CI_Tables$Awardee$PI_Sex.df
df_reviewer <- CI_Tables$Reviewer$Sex.df
df_applicant <- CI_Tables$Applicant$PI_Sex.df


df_panelist$Source <- "Panelists"
df_applicant$Source <- "Applicants (PI)"
df_awardee$Source <- "Awardees (PI)"
df_reviewer$Source <- "Reviewers"


# Combine dataframes
df_combined <- rbind(df_panelist, df_awardee, df_reviewer, df_applicant)
df_combined <- df_combined %>% filter(Sex == "Female")

df_combined <- df_combined %>% mutate(
  Year = substr(Year, 1, 4) 
)




female_proportions <- ggplot(df_combined, aes(x = Year, y = Proportion*100, color = Source)) +
  geom_point(position = position_dodge(width = 0.2), size = 3) +
  geom_errorbar(aes(ymin = BasicBootCI_L*100, ymax = BasicBootCI_U*100), width = 0.5, size = 0.8, alpha = 0.5, position = position_dodge(width = 0.2)) +
  labs(x = "Financial Year (Beginning)", 
       y = "Proportion (%)", 
       title = "Comparing EPSRC Population Female Proportions\nAcross Roles, by Year",
       color = "Population") +
  scale_color_manual(
    values = c(
        "Panelists" = RSSred,
        "Applicants (PI)" = RSSblue,
        "Awardees (PI)" = RSSgreen,
        "Reviewers" = RSSgold
    )
  ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1),
                     breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank()   
  )




print(female_proportions)
# Save the plot
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 02"

ggsave(filename = paste0(figure_folder_path, "/female_proportions_comparison.png"), 
       plot = female_proportions, width = 14, height = 12, dpi = 600, bg = "white")

```


## Ethnicity

#### Comparing All EPSRC Population Proportions Across Roles, by Year, by Ethnic Minority Proportion
```{r}

df_panelist <- CI_Tables$Panelist$Eth.df
df_awardee <- CI_Tables$Awardee$PI_Eth.df
df_reviewer <- CI_Tables$Reviewer$Eth.df
df_applicant <- CI_Tables$Applicant$PI_Eth.df


df_panelist$Source <- "Panelists"
df_applicant$Source <- "Applicants (PI)"
df_awardee$Source <- "Awardees (PI)"
df_reviewer$Source <- "Reviewers"


# Combine dataframes
df_combined <- rbind(df_panelist, df_awardee, df_reviewer, df_applicant)
df_combined <- df_combined %>% filter(EthnicityBinary == "Ethnic Minority")

df_combined <- df_combined %>% mutate(
  Year = substr(Year, 1, 4) 
)




ethnicity_proportions <- ggplot(df_combined, aes(x = Year, y = Proportion*100, color = Source)) +
  geom_point(position = position_dodge(width = 0.2), size = 3) +
  geom_errorbar(aes(ymin = BasicBootCI_L*100, ymax = BasicBootCI_U*100), width = 0.5, size = 0.8, alpha = 0.5, position = position_dodge(width = 0.2)) +
  labs(x = "Financial Year (Beginning)", 
       y = "Proportion (%)", 
       title = "Comparing EPSRC Population Ethnic Minority Proportions\nAcross Roles, by Year",
       color = "Population") +
  scale_color_manual(
    values = c(
        "Panelists" = RSSred,
        "Applicants (PI)" = RSSblue,
        "Awardees (PI)" = RSSgreen,
        "Reviewers" = RSSgold
    )
  ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1),
                     breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank()   
  )



print(ethnicity_proportions)
# Save the plot
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 02"

ggsave(filename = paste0(figure_folder_path, "/ethnicity_proportions_comparison.png"), 
       plot = ethnicity_proportions, width = 14, height = 12, dpi = 600, bg = "white")

```


## Nationality

#### Comparing All EPSRC Population Proportions Across Roles, by Year, by Non-UK Proportion
```{r}

df_panelist <- CI_Tables$Panelist$Nat.df
df_awardee <- CI_Tables$Awardee$PI_Nat.df
df_reviewer <- CI_Tables$Reviewer$Nat.df
df_applicant <- CI_Tables$Applicant$PI_Nat.df


df_panelist$Source <- "Panelists"
df_applicant$Source <- "Applicants (PI)"
df_awardee$Source <- "Awardees (PI)"
df_reviewer$Source <- "Reviewers"


# Combine dataframes
df_combined <- rbind(df_panelist, df_awardee, df_reviewer, df_applicant)
df_combined <- df_combined %>% filter(NationalityBinary == "Non UK")

df_combined <- df_combined %>% mutate(
  Year = substr(Year, 1, 4) 
)




nationality_proportions <- ggplot(df_combined, aes(x = Year, y = Proportion*100, color = Source)) +
  geom_point(position = position_dodge(width = 0.2), size = 3) +
  geom_errorbar(aes(ymin = BasicBootCI_L*100, ymax = BasicBootCI_U*100), width = 0.5, size = 0.8, alpha = 0.5, position = position_dodge(width = 0.2)) +
  labs(x = "Financial Year (Beginning)", 
       y = "Proportion (%)", 
       title = "Comparing EPSRC Population Non-UK Proportions\nAcross Roles, by Year",
       color = "Population") +
  scale_color_manual(
    values = c(
        "Panelists" = RSSred,
        "Applicants (PI)" = RSSblue,
        "Awardees (PI)" = RSSgreen,
        "Reviewers" = RSSgold
    )
  ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1),
                     breaks = seq(0, 40, by = 10),
                     limits = c(0, 40)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank()   
  )



print(nationality_proportions)
# Save the plot
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 02"

ggsave(filename = paste0(figure_folder_path, "/nationality_proportions_comparison.png"), 
       plot = nationality_proportions, width = 14, height = 12, dpi = 600, bg = "white")

```


## Disability

#### Comparing All EPSRC Population Proportions Across Roles, by Year, by Disabled Proportion
```{r}

df_panelist <- CI_Tables$Panelist$Dis.df
df_awardee <- CI_Tables$Awardee$PI_Dis.df
df_reviewer <- CI_Tables$Reviewer$Dis.df
df_applicant <- CI_Tables$Applicant$PI_Dis.df


df_panelist$Source <- "Panelists"
df_applicant$Source <- "Applicants (PI)"
df_awardee$Source <- "Awardees (PI)"
df_reviewer$Source <- "Reviewers"


# Combine dataframes
df_combined <- rbind(df_panelist, df_awardee, df_reviewer, df_applicant)
df_combined <- df_combined %>% filter(Disability == "Known Disability")

df_combined <- df_combined %>% mutate(
  Year = substr(Year, 1, 4) 
)




disability_proportions <- ggplot(df_combined, aes(x = Year, y = Proportion*100, color = Source)) +
  geom_point(position = position_dodge(width = 0.2), size = 3) +
  geom_errorbar(aes(ymin = BasicBootCI_L*100, ymax = BasicBootCI_U*100), width = 0.5, size = 0.8, alpha = 0.5, position = position_dodge(width = 0.2)) +
  labs(x = "Financial Year (Beginning)", 
       y = "Proportion (%)", 
       title = "Comparing EPSRC Population Proportions with\nKnown Disability Across Roles, by Year",
       color = "Population") +
  scale_color_manual(
    values = c(
        "Panelists" = RSSred,
        "Applicants (PI)" = RSSblue,
        "Awardees (PI)" = RSSgreen,
        "Reviewers" = RSSgold
    )
  ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1),
                     breaks = seq(0, 10, by = 2),
                     limits = c(0, 10)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank()   
  )



print(disability_proportions)
# Save the plot
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 02"

ggsave(filename = paste0(figure_folder_path, "/disability_proportions_comparison.png"), 
       plot = disability_proportions, width = 14, height = 12, dpi = 600, bg = "white")

```


## Age

#### Comparing All EPSRC Population Proportions Across Roles, by Year, by Age Group Proportion
```{r}

df_panelist <- CI_Tables$Panelist$Age.df
df_awardee <- CI_Tables$Awardee$PI_Age.df
df_reviewer <- CI_Tables$Reviewer$Age.df
df_applicant <- CI_Tables$Applicant$PI_Age.df


df_panelist$Source <- "Panelists"
df_applicant$Source <- "Applicants (PI)"
df_awardee$Source <- "Awardees (PI)"
df_reviewer$Source <- "Reviewers"


# Combine dataframes
df_combined <- rbind(df_panelist, df_awardee, df_reviewer, df_applicant)

df_combined <- df_combined %>% mutate(
  Year = substr(Year, 1, 4) 
)




age_proportions <- ggplot(df_combined %>% filter(AgeRange != "Unknown"), aes(x = Year, y = Proportion * 100, color = Source)) +
  geom_point(position = position_dodge(width = 0.2), size = 3) +
  geom_errorbar(aes(ymin = BasicBootCI_L*100, ymax = BasicBootCI_U*100), width = 0.5, size = 0.8, alpha = 0.5, position = position_dodge(width = 0.2)) +
  facet_wrap(~AgeRange, ncol = 3) +
  labs(x = "Financial Year (Beginning)", 
       y = "Proportion (%)", 
       title = "Comparing EPSRC Population Proportions by Age\nAcross Roles, by Year",
       color = "Population") +
  scale_color_manual(
    values = c(
        "Panelists" = RSSred,
        "Applicants (PI)" = RSSblue,
        "Awardees (PI)" = RSSgreen,
        "Reviewers" = RSSgold
    )
  ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1),
                     breaks = seq(0, 50, by = 10),
                     limits = c(0, 50)) +
  theme(
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"),
    plot.title.position = "plot",
    axis.title.x = element_text(size = 20, margin = margin(t = 15), face = "bold"),
    axis.title.y = element_text(size = 20, margin = margin(r = 10), face = "bold"),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10),
    axis.text.x = element_text(size = 16, angle = 90, face = "bold"),
    axis.text.y = element_text(size = 16, face = "bold"),
    legend.position = "top",
    legend.text = element_text(size = 16, face = "bold"),
    legend.title = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),  # Set panel background to white
    plot.background = element_rect(fill = "white", color = NA),   # Set plot background to white
    #panel.grid.major = element_line(color = "gray90"),            # Optional: adjust grid color
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 16, face = "bold")
  )



print(age_proportions)
# Save the plot
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 02"

ggsave(filename = paste0(figure_folder_path, "/age_proportions_comparison.png"), 
       plot = age_proportions, width = 14, height = 12, dpi = 600, bg = "white")

```