---
title: "Bias in EPSRC Peer Review - Checking Outcomes Against Missing and Not Disclosed Demographics"
author: 'Contributors: Ged Hiscoke & Dakota Langhals, Royal Statistical Society'
date: "Version: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    code_folding: hide
    toc_float: yes
  pdf_document: default
  word_document:
    reference_docx: "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/r_style_template.docx"
    highlight: null
subtitle: 'Status: Final'
---

```{r setup, include=FALSE}

##Universal Document settings
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12, warning = FALSE, messages = FALSE)

## Libraries
library(boot)
library(tidyverse)
library(kableExtra)
library(scales)
library(tibble)
library(purrr)
library(nortest)



## Directory for saving local high resolution copies of plots etc
data_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Datasets/"
note_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Notes/"
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Images/"
table_folder_path   <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/Tables/"

## Set key file names
raw_data_file       <- "PROJ2023.002-DATA001.xlsx"
BaselineEnvironment <- "PROJ2023.002 Environment - BASELINE.RData"
CleanedEnvironment  <- "PROJ2023.002 Environment - CLEANED.RData"
Sample20Environment <- "PROJ2023.002 Environment - SAMPLE20.RData"
Sample80Environment <- "PROJ2023.002 Environment - SAMPLE80.RData"
cleanedExcelFile    <- "PROJ2023.002-DATA002.xlsx"
WorkingEnvironment  <- "PROJ2023.002 Environment - WORKING.RData"
UKLFdata            <- "PROJ2023.002-DATA004.xlsx"


## Load needed data
load(file = file.path(data_folder_path, WorkingEnvironment))

## Define the RSS Colour palette

RSSgold       <- "#D3A41B" # Female
RSSblue       <- "#004B6D" # Male
RSSred        <- "#cc0000" # Not Disclosed
RSStaupe      <- "#a6a086" # Unknown
RSSgreen      <- "#009933" # White
RSSpurple     <- "#7879b9" # Non UK
RSSdark       <- "#393f4d" # UK
RSSlightblue  <- "#6db4c0" # Ethnic Minority
RSSseablue    <- "#287fc3" # No Known Disability
RSSorange     <- "#d86027"
RSSlightgrey  <- "#bdbdbd"
RSSmidgrey    <- "#939393"
RSSdarkkgrey  <- "#7e7e7e"
RSSjungle     <- "#1f3d0c"
RSSmint       <- "#5C9090" # Known Disability
RSSherb       <- "#689E88"
RSSbright     <- "#8D31A4"
RSSlime       <- "#7DCF7A"
RSSviolet     <- "#614182"

##Variable specific colours
col_male      <- "#004b6d"
col_female    <- "#a6700a"
col_uk        <- "#393f4d"
col_nonuk     <- "#614182"
col_dis       <- "#005b49"
col_nodis     <- "#8e8e00"
col_white     <- "#7f3c21"
col_ethmin    <- "#3a3877"
col_unknown   <- "#bdbdbd"
col_notdisclosed  <- "#640106"
col_25andunder <- "#b6d6b3"
col_2635 <- "#87bd00"
col_3645 <- "#299907"
col_4655 <- "#146f03"
col_5665 <- "#054907"
col_66 <- "#0f2808"
col_white <- "#a7a7c1"
col_chinese <- "#74749d"
col_black <- "#1e3399"
col_mixed <- "#090771"
col_asian <- "#37163e"
col_other <- "#59095f"
col_eth7 <- "#34086e"
```

# {.tabset}

## Data Preparation  
```{r}
## This sets up application-level dataframe

## 1.1 Select columns from Grants.df
Grants_selected.df <- EPSRC$Grants.df %>%
  select(AnonPROJ,
         GrantCategory,
         GrantOutcome,
         AdjustedApplicationValue
  )


## 1.2 Now select additional columns from PrincipalInvestigators.df
PI_details.df <- EPSRC$PI.df %>%
  select(
    AnonPROJ,
    AnonINV,
    Sex,
    AgeRange,
    Disability,
    EthnicGroup,
    EthnicityBinary,
    NationalityBinary,
    Role
  )


## 1.3 Merge with PrincipalInvestigators.df
Q0Grants.df <- left_join(PI_details.df,
                 Grants_selected.df ,
                 by = "AnonPROJ"
)


## 1.4 Cleanup
rm(PI_details.df, 
   Grants_selected.df
)

## Add a binary Outcome variable for Funded/Unfunded
Q0Grants.df$BinaryOutcome <- ifelse(Q0Grants.df$GrantOutcome == "Funded", "Funded", "Unfunded")

## Create factors that identify unknown, not disclosed, and disclosed for each demographic variable
Q0Grants.df <- Q0Grants.df %>% 
  mutate(
    Sex = ifelse(Sex == "Unknown", "Unknown", ifelse(Sex == "Not Disclosed", "Not Disclosed", "Disclosed")),
    Ethnicity = ifelse(EthnicityBinary == "Unknown", "Unknown", ifelse(EthnicityBinary == "Not Disclosed", "Not Disclosed", "Disclosed")),
    AgeRange = ifelse(AgeRange == "Unknown", "Unknown", ifelse(AgeRange == "Not Disclosed", "Not Disclosed", "Disclosed")),
    Nationality = ifelse(NationalityBinary == "Unknown", "Unknown", ifelse(NationalityBinary == "Not Disclosed", "Not Disclosed", "Disclosed")),
    Disability = ifelse(Disability == "Unknown", "Unknown", ifelse(Disability == "Not Disclosed", "Not Disclosed", "Disclosed")),
    All = if_else(
      if_all(c("Sex", "Ethnicity", "AgeRange", "Nationality", "Disability"), ~ . == "Disclosed"),
      "All Disclosed",
      "Not All Disclosed"
    )
  )

```

## Comparison of Binary Outcome for Rows with Complete and Incomplete Demographic Information, Application Level  

```{r, rows.print = 25}
# Calculate counts for BinaryOutcome plot
binary_counts <- Q0Grants.df %>%
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  group_by(Variable, Level, BinaryOutcome) %>%
  summarise(count = n(), .groups = 'drop')

binary_counts <- binary_counts %>% mutate(count = round(count / 10) * 10) ## For data suppression purposes
```

```{r}
# For BinaryOutcome
binary_plot <- Q0Grants.df %>% 
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  ggplot(aes(x = Level, fill = as.factor(BinaryOutcome))) +
  geom_bar(position = "fill") +
  geom_text(data = binary_counts, aes(x = Level, y = 1, label = count), 
            position = position_fill(vjust = 0.75), size = 5, color = "white") +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "BinaryOutcome by Levels of Demographic Variables",
       x = "Level",
       y = "Proportion",
       fill = "Binary Outcome") +
  scale_fill_manual(values = c("Funded" = RSSblue, "Unfunded" = RSSred)) +
  theme(plot.title = element_blank(), 
        axis.title.x = element_blank(), # Add space above x-axis title
        axis.title.y = element_text(size = 24, margin = margin(r = 15), face = "bold"), # Add space to the right of y-axis title
        axis.text.x = element_text(size = 16, angle = 25, vjust = 0.5), # Make axis easier to read
        axis.text.y = element_text(size = 16, face = "bold"),  # Make axis easier to read
        strip.text = element_text(size = 22, face = "bold"),
        strip.background = element_rect(fill = "white"),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = "bold"),
        legend.position = "top"
        )


binary_plot
```

## Comparison of Adjusted Award Value below £2.5m for Rows with Complete and Incomplete Demographic Information, Application Level  

```{r, rows.print = 25}

# Calculate counts for continuous plots, applying relevant filters
continuous_counts <- Q0Grants.df %>% filter(AdjustedApplicationValue < 2.5 & BinaryOutcome == "Funded") %>%
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  group_by(Variable, Level) %>%
  summarise(count = n(), .groups = 'drop')

continuous_counts <- continuous_counts %>% mutate(count = round(count / 10) * 10) ## For data suppression purposes


# For AdjustedApplicationValue
continuous_plot1 <- Q0Grants.df %>% filter(AdjustedApplicationValue < 2.5 & BinaryOutcome == "Funded") %>% 
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  ggplot(aes(x = Level, y = AdjustedApplicationValue)) +
  geom_violin() +
  geom_text(data = continuous_counts, aes(x = Level, y = 2.5, label = count), 
            position = position_fill(vjust = 1), size = 5, color = RSSred) +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "AdjustedApplicationValue below £2.5m by Levels of Demographic Variables",
       x = "Level",
       y = "Adjusted Application Value") +
  theme(plot.title = element_blank(), 
        axis.title.x = element_blank(), # Add space above x-axis title
        axis.title.y = element_text(size = 32, margin = margin(r = 15), face = "bold"), # Add space to the right of y-axis title
        axis.text.x = element_text(size = 22, face = "bold"), # Make axis easier to read
        axis.text.y = element_text(size = 22, face = "bold"),  # Make axis easier to read
        strip.text = element_text(size = 22, face = "bold"),
        strip.background = element_rect(fill = "white")
        )



demographic_vars <- c("Sex", "Ethnicity", "AgeRange", "Nationality", "Disability", "All")
# Loop through each demographic variable and calculate summary statistics
summary_stats <- Q0Grants.df %>% filter(AdjustedApplicationValue < 2.5 & BinaryOutcome == "Funded") %>%
  pivot_longer(cols = all_of(demographic_vars), 
               names_to = "Variable", values_to = "Level") %>%
  filter(Level != "Unknown" & !is.na(Level)) %>%
  group_by(Variable, Level) %>%
  summarise(
    count = n(),
    mean = mean(AdjustedApplicationValue, na.rm = TRUE),
    median = median(AdjustedApplicationValue, na.rm = TRUE),
    min = min(AdjustedApplicationValue, na.rm = TRUE),
    max = max(AdjustedApplicationValue, na.rm = TRUE),
    sd = sd(AdjustedApplicationValue, na.rm = TRUE)
  )

# Display the results
summary_stats <- summary_stats %>% mutate(count = round(count / 10) * 10) ## For data suppression purposes

print(summary_stats)


continuous_plot1

```


## Comparison of Adjusted Award Value above £2.5m for Rows with Complete and Incomplete Demographic Information, Application Level  

```{r, rows.print = 25}
# Calculate counts for continuous plots, applying relevant filter
continuous_counts <- Q0Grants.df %>% filter(AdjustedApplicationValue >= 2.5 & BinaryOutcome == "Funded") %>%
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  group_by(Variable, Level) %>%
  summarise(count = n(), .groups = 'drop')

continuous_counts <- continuous_counts %>% mutate(count = round(count / 10) * 10) ## For data suppression purposes




continuous_plot2 <- Q0Grants.df %>% filter(AdjustedApplicationValue >= 2.5 & BinaryOutcome == "Funded") %>% 
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  ggplot(aes(x = Level, y = AdjustedApplicationValue)) +
  geom_violin() +
  geom_text(data = continuous_counts, aes(x = Level, y = 2.5, label = count), 
            position = position_fill(vjust = -0.5), size = 3, color = "black") +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "AdjustedApplicationValue above £2.5m by Levels of Demographic Variables",
       x = "Level",
       y = "Adjusted Application Value") +
  theme(
      plot.title = element_blank(), 
      axis.title.x = element_blank(), # Add space above x-axis title
      axis.title.y = element_text(size = 32, margin = margin(r = 15), face = "bold"), # Add space to the right of y-axis title
      plot.margin = margin(t = 20, r = 10, b = 10, l = 10), # Add extra space around the plot
      axis.text.x = element_text(size = 16, face = "bold", angle = 90), # Make axis easier to read
      axis.text.y = element_text(size = 22, face = "bold"),  # Make axis easier to read
      strip.text = element_text(size = 16, face = "italic"),
      legend.position = "top",
      legend.text = element_text(size = 16, face = "bold")
  )


demographic_vars <- c("Sex", "Ethnicity", "AgeRange", "Nationality", "Disability", "All")
# Loop through each demographic variable and calculate summary statistics
summary_stats <- Q0Grants.df %>% filter(AdjustedApplicationValue >= 2.5 & BinaryOutcome == "Funded") %>%
  pivot_longer(cols = all_of(demographic_vars), 
               names_to = "Variable", values_to = "Level") %>%
  filter(Level != "Unknown" & !is.na(Level)) %>%
  group_by(Variable, Level) %>%
  summarise(
    count = n(),
    mean = mean(AdjustedApplicationValue, na.rm = TRUE),
    median = median(AdjustedApplicationValue, na.rm = TRUE),
    min = min(AdjustedApplicationValue, na.rm = TRUE),
    max = max(AdjustedApplicationValue, na.rm = TRUE),
    sd = sd(AdjustedApplicationValue, na.rm = TRUE)
  )

# Display the results
summary_stats <- summary_stats %>% mutate(count = round(count / 10) * 10) ## For data suppression purposes

print(summary_stats)



continuous_plot2


```


## Comparison of Top Quartile Panel Ranking for Rows with Complete and Incomplete Demographic Information, Application Level  

```{r}

#----Identify the meeting info----------------------------------------------- 
# Make sure the ranking is numberic 
EPSRC$GrantMeeting.df$MeetingOrder <- as.numeric(EPSRC$GrantMeeting.df$MeetingOrder) 

# Create temp.df with rows of the final meeting 
temp.df <- EPSRC$GrantMeeting.df %>% 
  group_by(AnonPROJ) %>% 
  filter(MeetingOrder == max(MeetingOrder, na.rm = TRUE)) %>% 
  ungroup() 

# fill in NAs in not-fundable 
temp.df <- temp.df %>% 
  mutate(FundableByScore = ifelse(is.na(FundableByScore), 
                                  ifelse(Decision == "Not Fundable", FALSE, TRUE), 
                                  FundableByScore)) 

# Remove applications with no ranking.  
temp.df <- temp.df %>% 
  filter(!is.na(RankOrder)) 

#----SMerge Quantiles into the main dataset------------------------------------ 
Q0Grants.df <- Q0Grants.df%>% 
  left_join(temp.df %>%  
              select(AnonPROJ, RankOrder, FundableByScore, AnonMTG),  
            by = "AnonPROJ") 

# Adjust FundableByScore based on BinaryOutcome 
Q0Grants.df <- Q0Grants.df %>% 
  mutate(FundableByScore = if_else(BinaryOutcome == "Funded" & FundableByScore == FALSE, TRUE, FundableByScore)) 

calculate_quartiles <- function(rank_order, fundable_by_score) { 
  # Adjust for NA or 0 in rank_order or NA in fundable_by_score 
  # Treat 0 in RankOrder as equivalent to NA for the purpose of ranking 
  rank_order[rank_order == 0] <- NA 
  # If all fundable_by_score are NA, but rank_order has valid numbers, proceed as if they are fundable 
  if (all(is.na(fundable_by_score)) && any(!is.na(rank_order))) { 
    fundable_by_score[is.na(fundable_by_score)] <- TRUE 
  } 

  # If fundable_by_score is explicitly FALSE, treat all as "Unfundable" 
  if (all(fundable_by_score == FALSE, na.rm = TRUE)) { 
    return(rep("Unfundable", length(rank_order))) 
  } else { 
    # Check for unique valid values in rank_order 
    valid_ranks <- unique(rank_order[!is.na(rank_order)]) 
    if (length(valid_ranks) < 2) { 
      return(rep("Unranked", length(rank_order))) 
    } else { 
      # Calculate quartiles for valid rank_order values, ensure breaks are unique 
      quartiles <- quantile(rank_order[!is.na(rank_order)], probs = c(0.25, 0.5, 0.75), na.rm = TRUE) 
      if (length(unique(quartiles)) < 3) {  # Check if quartile values are not unique 
        return(rep("Unranked", length(rank_order))) 
      } 
      breaks = c(-Inf, quartiles, Inf) 
      labels = c("1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile") 
      quartile_labels <- cut(rank_order, breaks = breaks, labels = labels, include.lowest = TRUE) 
      return(as.character(quartile_labels)) 
    } 
  } 
} 

Q0Grants.df <- Q0Grants.df %>% 
  group_by(AnonMTG, FundableByScore) %>% 
  mutate(QuartileInfo = calculate_quartiles(RankOrder, FundableByScore)) %>% 
  ungroup() 

#----Clean up dataset---------------------------------------------------------- 
# Replace NA in QuartileInfo with "Unranked" 
Q0Grants.df <- Q0Grants.df %>% 
  mutate(QuartileInfo = if_else(is.na(QuartileInfo), "Unranked", QuartileInfo)) 

Q0Grants.df$QuartileInfo <- factor(Q0Grants.df$QuartileInfo, 
                                    levels = c("1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile","Unfundable","Unranked"), 
                                    ordered = TRUE) 

# Remove temp.df from the environment 
rm(temp.df) 


# New binary outcome for top ranking = *TopQuartile*:
Q0Grants.df  <- Q0Grants.df %>% 
  mutate(TopQuartile = if_else(QuartileInfo == "1st Quartile", QuartileInfo, "Not 1st Quartile")) 

```


```{r, rows.print = 25}
# Calculate counts for TopQuartile plot
top_quartile_counts <- Q0Grants.df %>%
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  group_by(Variable, Level, TopQuartile) %>%
  summarise(count = n(), .groups = 'drop')

top_quartile_counts <- top_quartile_counts %>% mutate(count = round(count / 10) * 10) ## For data suppression purposes


# For TopQuartile
top_quartile_plot <- Q0Grants.df %>% 
  pivot_longer(cols = c(Sex, Ethnicity, AgeRange, Nationality, Disability, All),
               names_to = "Variable", values_to = "Level") %>%
  ggplot(aes(x = Level, fill = as.factor(TopQuartile))) +
  geom_bar(position = "fill") +
  geom_text(data = top_quartile_counts, aes(x = Level, y = 1, label = count), 
            position = position_fill(vjust = 0.75), size = 5, color = "white") +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "TopQuartile by Levels of Demographic Variables",
       x = "Level",
       y = "Proportion",
       fill = "Top Quartile Ranking") +
  scale_fill_manual(values = c("1st Quartile" = RSSblue, "Not 1st Quartile" = RSSred)) +
  theme(plot.title = element_blank(), 
        axis.title.x = element_blank(), # Add space above x-axis title
        axis.title.y = element_text(size = 24, margin = margin(r = 15), face = "bold"), # Add space to the right of y-axis title
        axis.text.x = element_text(size = 16, angle = 25, vjust = 0.5), # Make axis easier to read
        axis.text.y = element_text(size = 16, face = "bold"),  # Make axis easier to read
        strip.text = element_text(size = 22, face = "bold"),
        strip.background = element_rect(fill = "white"),
        legend.text = element_text(size = 22),
        legend.title = element_text(size = 22, face = "bold"),
        legend.position = "top"
        )


top_quartile_plot


```
