---
title: "Bias in EPSRC Peer Review - Research Question 17"
author: 'Contributors: Ged Hiscoke & Dakota Langhals, Royal Statistical Society'
date: "Version: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S')`"
output:
  html_document:
    code_folding: hide
    toc_float: yes
  pdf_document: default
  word_document:
    reference_docx: "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Draft Outputs/r_style_template.docx"
    highlight: null
subtitle: 'Status: Final'
---

```{r setup, echo = FALSE, include=FALSE}

##Universal Document settings
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 12, warning = FALSE, message = FALSE)


## Libraries
library(dplyr)      # For tidy coding
library(kableExtra) # For tables
library(ggplot2)    # For plotting
library(patchwork)  # For plot displays
library(tidyr)      # For tidy coding
library(ggforce)    # For tidy labelling on plots 
library(gridExtra)  # For display of joint plots
library(knitr)
library(openxlsx)   # for saving tables as excel files
library(scales)     # For commas in plot labels
library(DescTools)
library(boot)
library(stringr)





## Set key file locations
## Directory for saving local high resolution copies of plots etc
data_folder_path    <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/PROJ2023.002 - Data folder DL Pickup/Datasets/"
figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"




## Set key file names
WorkingEnvironment  <- "PROJ2023.002 Environment - WORKING.RData"


## Load needed data
load(file = file.path(data_folder_path, WorkingEnvironment))



## Define the RSS colour palette

RRSSgold      <- "#D3A41B" 
RSSblue       <- "#004B6D" 
RSSred        <- "#cc0000" 
RSStaupe      <- "#a6a086" 
RSSgreen      <- "#009933" 
RSSpurple     <- "#7879b9" 
RSSdark       <- "#393f4d" 
RSSlightblue    <- "#6db4c0" 
RSSseablue       <- "#287fc3" 
RSSorange     <- "#d86027"
RSSlightgrey  <- "#bdbdbd"
RSSmidgrey    <- "#939393"
RSSdarkkgrey  <- "#7e7e7e"
RSSjungle     <- "#1f3d0c"
RSSmint       <- "#5C9090" 
RSSherb       <- "#689E88"
RSSbright     <- "#8D31A4"
RSSlime       <- "#7DCF7A"
RSSviolet     <- "#614182"


##Variable specific colours
col_male      <- "#004b6d"
col_female    <- "#a6700a"
col_uk        <- "#393f4d"
col_nonuk     <- "#614182"
col_dis       <- "#005b49"
col_nodis     <- "#8e8e00"
col_white     <- "#7f3c21"
col_ethmin    <- "#3a3877"
col_unknown   <- "#bdbdbd"
col_notdisclosed  <- "#640106"
col_age1 <- "#b6d6b3"
col_age2 <- "#87bd00"
col_age3 <- "#299907"
col_age4 <- "#146f03"
col_age5 <- "#054907"
col_age6 <- "#0f2808"
col_eth1 <- "#a7a7c1"
col_eth2 <- "#74749d"
col_eth3 <- "#1e3399"
col_eth4 <- "#090771"
col_eth5 <- "#37163e"
col_eth6 <- "#59095f"
col_eth7 <- "#34086e"
```


# {.tabset}

## Data Preparation

```{r dataprep}

## Select the main data for the note, and then filter to unique individuals per year x Driver
df <- EPSRC$Investigators.df %>% select(
   AnonINV,
   Driver,
   Sex, 
   EthnicityBinary, 
   NationalityBinary,
   Disability,
   AgeRange,
   FYDecisionDate
   )


Q17data.df <- df %>%                     
  group_by(FYDecisionDate, Driver) %>%
  distinct(AnonINV, .keep_all = TRUE) %>%    # Keep only one unique AnonINV
  ungroup()


# Collapsing age to align with ATI and make interactions more easily interpretable
Q17data.df <- Q17data.df %>% 
  mutate(AgeRangeCollapsed = case_when(AgeRange == "25 and Under" ~ "35_and_Under", 
                                       AgeRange == "26-35" ~ "35_and_Under",
                                       AgeRange == "36-45" ~ "36-55",
                                       AgeRange == "46-55" ~ "36-55",
                                       AgeRange == "56-65" ~ "56_and_Over",
                                       AgeRange == "66+" ~  "56_and_Over"))

Q17data.df$FYDecisionDate <- substr(Q17data.df$FYDecisionDate, 1, 4)

# Intersectional demographics
Q17data.df$Sex_Ethnicity <- interaction(Q17data.df$Sex, Q17data.df$EthnicityBinary)
Q17data.df$Sex_Age <- interaction(Q17data.df$Sex, Q17data.df$AgeRangeCollapsed)
Q17data.df$Ethnicity_Age <- interaction(Q17data.df$EthnicityBinary, Q17data.df$AgeRangeCollapsed)

```

```{r confidence_interval_function}
##Function-----------------------------------------
# Function to calculate bootstrap CI for proportions
boot_ci <- function(data, indices) {
  sampled_data <- data[indices]
  prop <- mean(sampled_data)
  return(prop)
}

# helper Function to compute bootstrapped confidence intervals
compute_ci <- function(data) {
  if (length(data) < 2) {
    return(data.frame(CI_Lower = NA, CI_Upper = NA))
  }
  
  boot_result <- boot(data, boot_ci, R = 1000)
  ci <- boot.ci(boot_result, type = "basic")
  return(data.frame(CI_Lower = ci$basic[4], CI_Upper = ci$basic[5]))
}

# Generalized function to compute counts, proportions, and CIs by group and demographic
compute_statistics <- function(df, group_col, demo_col, driver = "Uncertain") {
  df <- df %>%
    rename(Group = !!sym(group_col), Demographic = !!sym(demo_col))
  
  groups <- unique(df$Group)
  results_list <- list()
  
  for (group in groups) {
    group_df <- df %>% filter(Group == group)
    
    result <- group_df %>%
      group_by(Demographic) %>%
      summarise(
        Count = n(),
        Proportion = Count / nrow(group_df)
      ) %>%
      ungroup() %>%
      rowwise() %>%
      mutate(
        CI = list(compute_ci(group_df$Demographic == Demographic)),
        CI_L = CI$CI_Lower,
        CI_U = CI$CI_Upper
      ) %>%
      select(-CI) %>%
      ungroup() %>%
      mutate(Year = group)
    
    results_list[[group]] <- result
  }
  
  results_df <- bind_rows(results_list)
  results_df <- bind_rows(results_list) %>%
    mutate(Group = demo_col)
  
  results_df$Driver <- driver
  return(results_df)
}



```

```{r table_rounding_function}
# Function to round columns as specified
round_columns <- function(df) {
  df <- df %>%
    mutate(
      Count = ifelse(Count < 6, "\u22645", round(Count / 5) * 5),
      across(where(is.numeric) & !c(Count), ~ ifelse(Count == "\u22645", "~", round(., 1)))
    )
  return(df)
}

```

```{r}
## Defining defaults for plots

update_geom_defaults("line", list(alpha = 0.5))

update_geom_defaults("errorbar", list(size = 1.2, width = 0.5, alpha = 0.5))

update_geom_defaults("point", list(size = 3))


theme_my_defaults <- function() {
  theme(
    text = element_text(family = "Frutiger LT Std"),
    plot.title = element_text(size = 24, hjust = 0.5, margin = margin(b = 15), face = "bold"), # Center title, add space below
    axis.title.x = element_text(size = 22, margin = margin(t = 15), face = "bold", colour = "black"), # Add space above x-axis title
    axis.title.y = element_text(size = 22, margin = margin(r = 15), face = "bold", colour = "black"), # Add space to the right of y-axis title
    # plot.margin = margin(t = 20, r = 10, b = 10, l = 10), # Add extra space around the plot
    axis.text.x = element_text(size = 16, face = "bold", angle = 90, colour = "black"), # Make axis easier to read
    axis.text.y = element_text(size = 16, face = "bold", colour = "black"),  # Make axis easier to read
    plot.title.position = "plot",
    panel.background = element_rect(fill = "white"),
    strip.text = element_text(size = 16, face = "italic"),
    legend.position = "top",
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18, face = "bold")
  ) 
}
```


## Tables: Proportions of Individuals Applying for Strategic and Responsive Awards by Demographics {.tabset}

#### Table 17.01: Proportions of Individuals Applying for Strategic and Responsive Awards by Demographics

```{r}
df <- Q17data.df%>% filter(Driver =="Strategic" & !is.na(AgeRangeCollapsed))
t3 <- compute_statistics(df %>% filter(Sex != "Unknown", Sex != "Not Disclosed"), "FYDecisionDate", "Sex", "Strategic")
t3 <- rbind(t3,compute_statistics(df %>% filter(EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed"), "FYDecisionDate", "EthnicityBinary", "Strategic"))
t3 <- rbind(t3,compute_statistics(df %>% filter(NationalityBinary != "Unknown", NationalityBinary != "Not Disclosed"), "FYDecisionDate", "NationalityBinary", "Strategic"))
t3 <- rbind(t3,compute_statistics(df %>% filter(Disability != "Unknown", Disability != "Not Disclosed"), "FYDecisionDate", "Disability", "Strategic"))
t3 <- rbind(t3,compute_statistics(df %>% filter(AgeRangeCollapsed != "Unknown", AgeRangeCollapsed != "Not Disclosed"), "FYDecisionDate", "AgeRangeCollapsed", "Strategic"))

df <- Q17data.df%>% filter(Driver =="Responsive" & !is.na(AgeRangeCollapsed))
t3 <- rbind(t3,compute_statistics(df %>% filter(Sex != "Unknown", Sex != "Not Disclosed"), "FYDecisionDate", "Sex", "Responsive"))
t3 <- rbind(t3,compute_statistics(df %>% filter(EthnicityBinary != "Unknown", EthnicityBinary != "Not Disclosed"), "FYDecisionDate", "EthnicityBinary", "Responsive"))
t3 <- rbind(t3,compute_statistics(df %>% filter(NationalityBinary != "Unknown", NationalityBinary != "Not Disclosed"), "FYDecisionDate", "NationalityBinary", "Responsive"))
t3 <- rbind(t3,compute_statistics(df %>% filter(Disability != "Unknown", Disability != "Not Disclosed"), "FYDecisionDate", "Disability", "Responsive"))
t3 <- rbind(t3,compute_statistics(df %>% filter(AgeRangeCollapsed != "Unknown", AgeRangeCollapsed != "Not Disclosed"), "FYDecisionDate", "AgeRangeCollapsed", "Responsive"))

table17.01 <- t3 %>% round_columns() # rounding to avoid revealing sensitive information.

kable(table17.01 %>% arrange(Group, Year, Driver))
```

<br>
#### Table 17.02: Proportions of Individuals Applying for Strategic and Responsive Awards by Intersectional Demographics
```{r}

df <- Q17data.df%>% filter(Driver =="Strategic" & !is.na(AgeRangeCollapsed))
intersections <- compute_statistics(df, "FYDecisionDate", "Sex_Ethnicity", "Strategic")
intersections <- rbind(intersections, compute_statistics(df, "FYDecisionDate", "Sex_Age", "Strategic"))
intersections <- rbind(intersections, compute_statistics(df, "FYDecisionDate", "Ethnicity_Age", "Strategic"))

df <- Q17data.df%>% filter(Driver =="Responsive" & !is.na(AgeRangeCollapsed))
intersections <- rbind(intersections, compute_statistics(df, "FYDecisionDate", "Sex_Ethnicity", "Responsive"))
intersections <- rbind(intersections, compute_statistics(df, "FYDecisionDate", "Sex_Age", "Responsive"))
intersections <- rbind(intersections, compute_statistics(df, "FYDecisionDate", "Ethnicity_Age", "Responsive"))

table17.02 <- intersections %>% 
  filter(!str_detect(Demographic, "Unknown|Not Disclosed")) %>% 
  round_columns() # rounding to avoid revealing sensitive information.

kable(table17.02 %>% arrange(Group, Year, Driver))
```

<br>

## Plots: One-dimensional Demographics

<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Disability 
``` {r}
plot_colors <-
  c(
    "Male"= col_male,      
    "Female"= col_female,    
    "UK"=col_uk,        
    "Non UK"=col_nonuk,     
    "Known Disability"=RSSred   ,    
    "No Known Disability"=RSSblue  ,   
    "White"=col_white  ,   
    "Ethnic Minority"=col_ethmin  ,  
    "Unknown"=col_unknown  , 
    "Not Disclosed"=col_notdisclosed  ,
    "35_and_Under"=col_25andunder ,
    "36-55"=col_2635 ,
    "56_and_Over"=col_3645 ,
    "White"=col_white, 
    "Chinese"=col_chinese, 
    "Black"=col_black, 
    "Mixed"=col_mixed, 
    "Asian"=col_asian, 
    "Other"=col_other
    
  )


t3$Year  <- factor(t3$Year, levels = c("2014",
                                       "2015",
                                       "2016",
                                       "2017",
                                       "2018",
                                       "2019",
                                       "2020",
                                       "2021",
                                       "2022"))

t3_filtered <- t3 %>%filter(Demographic != "Unknown" & 
                     Demographic != "Not Disclosed")


df <- t3_filtered %>%  filter(Group == "Disability")

df <- df %>% rename(Mode = Driver)

disability_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), size = 3, aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
                position = position_dodge(width = 0.5),
                size = 1) +
  scale_color_manual(values = plot_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = "Rates of Individuals Applying for Strategic and Responsive\nFunding Opportunities, by Disability",
       x = "Year",
       y = "Percentage") +
  theme_my_defaults()

disability_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/disability_plot.png"), 
       plot = disability_plot, width = 12, height = 12, dpi = 600, bg = "white")
```

<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Nationality
``` {r}
plot_colors <-
  c(
    "Male"= col_male,      
    "Female"= col_female,    
    "UK"=RSSblue,        
    "Non UK"=RSSred,     
    "White"=col_white  ,   
    "Ethnic Minority"=col_ethmin  ,  
    "Unknown"=col_unknown  , 
    "Not Disclosed"=col_notdisclosed  ,
    "35_and_Under"=col_25andunder ,
    "36-55"=col_2635 ,
    "56_and_Over"=col_3645 ,
    "White"=col_white, 
    "Chinese"=col_chinese, 
    "Black"=col_black, 
    "Mixed"=col_mixed, 
    "Asian"=col_asian, 
    "Other"=col_other
    
  )


df <- t3_filtered %>%  filter(Group == "NationalityBinary")

df <- df %>% rename(Mode = Driver)

nationality_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), size = 3, aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
                position = position_dodge(width = 0.5),
                size = 1) +
  scale_color_manual(values = plot_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = "Rates of Individuals Applying for Strategic and Responsive\nFunding Opportunities, by Nationality",
       x = "Year",
       y = "Percentage") +
  theme_my_defaults()

nationality_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/nationality_plot.png"), 
       plot = nationality_plot, width = 12, height = 12, dpi = 600, bg = "white")
```


<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Sex
``` {r}
plot_colors <-
  c(
    "Male"= RSSred,      
    "Female"= RSSblue,    
    "UK"=col_uk,        
    "Non UK"=col_nonuk,     
    "White"=col_white  ,   
    "Ethnic Minority"=col_ethmin  ,  
    "Unknow"=col_unknown  , 
    "Not Disclosed"=col_notdisclosed  ,
    "35_and_Under"=col_25andunder ,
    "36-55"=col_2635 ,
    "56_and_Over"=col_3645 ,
    "White"=col_white, 
    "Chinese"=col_chinese, 
    "Black"=col_black, 
    "Mixed"=col_mixed, 
    "Asian"=col_asian, 
    "Other"=col_other
    
  )


df <- t3_filtered %>%  filter(Group == "Sex")

df <- df %>% rename(Mode = Driver)

sex_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), size = 3, aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
                position = position_dodge(width = 0.5),
                size = 1) +
  scale_color_manual(values = plot_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = "Rates of Individuals Applying for Strategic and Responsive\nFunding Opportunities, by Sex",
       x = "Year",
       y = "Percentage") +
  theme_my_defaults()

sex_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/sex_plot.png"), 
       plot = sex_plot, width = 12, height = 12, dpi = 600, bg = "white")
```

<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Ethnicity
``` {r}
plot_colors <-
  c(
    "UK"=col_uk,        
    "Non UK"=col_nonuk,     
    "White"=RSSblue  ,   
    "Ethnic Minority"=RSSred  ,  
    "Unknown"=col_unknown  , 
    "Not Disclosed"=col_notdisclosed  ,
    "35_and_Under"=col_25andunder ,
    "36-55"=col_2635 ,
    "56_and_Over"=col_3645 ,
    "White"=col_white, 
    "Chinese"=col_chinese, 
    "Black"=col_black, 
    "Mixed"=col_mixed, 
    "Asian"=col_asian, 
    "Other"=col_other
)


df <- t3_filtered %>%  filter(Group == "EthnicityBinary")

df <- df %>% rename(Mode = Driver)

ethnicity_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), size = 3, aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
                position = position_dodge(width = 0.5),
                size = 1) +
  scale_color_manual(values = plot_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = "Rates of Individuals Applying for Strategic and Responsive\nFunding Opportunities, by Ethnicity",
       x = "Year",
       y = "Percentage") +
  theme_my_defaults()

ethnicity_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/ethnicity_plot.png"), 
       plot = ethnicity_plot, width = 12, height = 12, dpi = 600, bg = "white")
```


<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Age
``` {r}
plot_colors <-
  c(
    "UK"=col_uk,        
    "Non UK"=col_nonuk,     
    "Unknown"=col_unknown  , 
    "Not Disclosed"=col_notdisclosed  ,
    "35_and_Under"=RSSred ,
    "36-55"=RSSblue ,
    "56_and_Over"=RSSgold ,
    "White"=col_white, 
    "Chinese"=col_chinese, 
    "Black"=col_black, 
    "Mixed"=col_mixed, 
    "Asian"=col_asian, 
    "Other"=col_other
    
  )


df <- t3_filtered %>%  filter(Group == "AgeRangeCollapsed")

df <- df %>% rename(Mode = Driver)

age_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), size = 3, aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U), 
                position = position_dodge(width = 0.5),
                size = 1) +
  scale_color_manual(values = plot_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = "Rates of Individuals Applying for Strategic and Responsive\nFunding Opportunities, by Age Group",
       x = "Year",
       y = "Percentage") +
  theme_my_defaults()

age_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/age_plot.png"), 
       plot = age_plot, width = 12, height = 12, dpi = 600, bg = "white")
```



## Plots: Intersectional Demographics

```{r}
# Intersectional plots
intersections$Year  <- factor(intersections$Year, levels =
                                    c("2014",
                                       "2015",
                                       "2016",
                                       "2017",
                                       "2018",
                                       "2019",
                                       "2020",
                                       "2021",
                                       "2022"))

intersections_filtered <- intersections %>% filter(!str_detect(Demographic, "Unknown|Not Disclosed"))
```

<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Intersectional Sex and Ethnicity
```{r}

plot_colors <-
  c(
    "Male.White"= RSSblue,
    "Female.White"= RSSred,
    "Male.Ethnic Minority"=RSSgreen,
    "Female.Ethnic Minority"=RSSgold,
    "Female.35_and_Under"=col_dis   ,
    "Female.36-55"=col_nodis  ,
    "Female.56_and_Over"=col_white  ,
    "Male.35_and_Under"=col_ethmin  ,
    "Male.36-55"=col_unknown  ,
    "Male.56_and_Over"=col_notdisclosed  ,
    "White.35_and_Under"=col_25andunder ,
    "White.36-55"=col_2635 ,
    "White.56_and_Over"=col_3645 ,
    "Ethnic Minority.35_and_Under"=col_age2 ,
    "Ethnic Minority.36-55"=col_eth3 ,
    "Ethnic Minority.56_and_Over"=col_eth6
  )

df <- intersections_filtered %>% filter(Group == "Sex_Ethnicity")

df <- df %>% rename(Mode = Driver)

sex_ethnicity_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U),
                position = position_dodge(width = 0.5)) +
  scale_color_manual(values = plot_colors) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = element_blank(),
       x = "Proportion",
       y = "Year") +
  theme_my_defaults()

sex_ethnicity_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/sex_ethnicity_plot.png"), 
       plot = sex_ethnicity_plot, width = 12, height = 12, dpi = 600, bg = "white")
```


<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Intersectional Sex and Age
```{r}
plot_colors <-
  c(
    "Female.35_and_Under"=RSSred   ,
    "Female.36-55"=RSSgold,
    "Female.56_and_Over"=RSSbright  ,
    "Male.35_and_Under"=RSSblue  ,
    "Male.36-55"=RSSgreen  ,
    "Male.56_and_Over"=RSSlightblue  ,
    "White.35_and_Under"=col_25andunder ,
    "White.36-55"=col_2635 ,
    "White.56_and_Over"=col_3645 ,
    "Ethnic Minority.35_and_Under"=col_age2 ,
    "Ethnic Minority.36-55"=col_eth3 ,
    "Ethnic Minority.56_and_Over"=col_eth6
  )


df <- intersections_filtered %>% filter(Group == "Sex_Age")

df <- df %>% rename(Mode = Driver)

sex_age_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U),
                position = position_dodge(width = 0.5)) +
  scale_color_manual(values = plot_colors) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = element_blank(),
       x = "Proportion",
       y = "Year") +
  theme_my_defaults()

sex_age_plot


figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/sex_age_plot.png"), 
       plot = sex_age_plot, width = 12, height = 12, dpi = 600, bg = "white")
```

<br>
#### Proportions of Individuals Applying for Strategic and Responsive Awards by Intersectional Ethnicity and Age
```{r}
plot_colors <-
  c(
    "White.35_and_Under"=RSSred ,
    "White.36-55"=RSSgold ,
    "White.56_and_Over"=RSSbright ,
    "Ethnic Minority.35_and_Under"=RSSblue ,
    "Ethnic Minority.36-55"=RSSgreen ,
    "Ethnic Minority.56_and_Over"=RSSlightblue
  )

df <- intersections_filtered %>% filter(Group == "Ethnicity_Age")

df <- df %>% rename(Mode = Driver)

ethnicity_age_plot <- ggplot(df, aes(x = Year, y = Proportion, color = Demographic, group = interaction(Demographic, Mode))) +
  geom_line(position = position_dodge(width = 0.5), size = 1, aes(linetype = Mode)) +  # Add lines for trends
  geom_point(position = position_dodge(width = 0.5), aes(shape = Mode)) +  # Add points
  geom_errorbar(aes(ymin = CI_L, ymax = CI_U),
                position = position_dodge(width = 0.5)) +
  scale_color_manual(values = plot_colors) +
  scale_shape_manual(values = c("Responsive" = 16, "Strategic" = 17)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_linetype_manual(values = c("Responsive" = "solid", "Strategic" = "dashed")) +  # Different line types
  labs(title = element_blank(),
       x = "Proportion",
       y = "Year") +
  theme_my_defaults()

ethnicity_age_plot



figure_folder_path  <- "C:/Users/DakotaLanghals/OneDrive - Royal Statistical Society/EPSRC Grant Funding/Figures/Research Question 17"

ggsave(filename = paste0(figure_folder_path, "/ethnicity_age_plot.png"), 
       plot = ethnicity_age_plot, width = 12, height = 12, dpi = 600, bg = "white")
```



